---
title: "MatosAnalysis"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(latex2exp)
library(readxl)
library(egg)
library(plotly)
library(readr)
library(neonUtilities) #download and manage NEON data (in-house)
library(neonOS) #NEON OS data
library(terra) #general geological data
library(ggplot2)
library(colorspace)
library(viridis)
library(sf)
library(tigris)
library(leaflet)
library(maps)
library(crosstalk)
library(lubridate)
library(wsyn)
```

## Loading New Data Products

Need:

-   atmos vapor concentration

-   soil temperature

-   daily air temp

-   solar radiation

Have:

-   atmos isotope concentration

-   precipitation amount

-   precipitation isotope concentration

-   relative humidity

```{r}
# soil radiation
SoilRadiation <- loadByProduct(
  dpID = "DP1.00014.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE)
```

```{r}
#single aspirated air temp
SingleAspiratedAirTemp <- loadByProduct(
  dpID = "DP1.00002.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE)
```

```{r}
#triple aspirated air temp
TripleAspiratedAirTemp <- loadByProduct(
  dpID = "DP1.00003.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE)
```

```{r}
SoilTemperature2022 <- loadByProduct(
  dpID = "DP1.00041.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2022-12",
  package = "basic",
  check.size = FALSE
)
  
```

```{r}
SoilTemperature2023 <- loadByProduct(
  dpID = "DP1.00041.001",
  site = "ONAQ",
  startdate = "2023-01",
  enddate = "2023-12",
  package = "basic",
  check.size = FALSE
)
```

```{r}
SoilTemperature2024 <- loadByProduct(
  dpID = "DP1.00041.001",
  site = "ONAQ",
  startdate = "2024-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE
)
```

```{r}
write_rds(SoilTemperature2022$ST_30_minute, "SoilTemperature2022.rds")
write_rds(SoilTemperature2023$ST_30_minute, "SoilTemperature2023.rds")
write_rds(SoilTemperature2024$ST_30_minute, "SoilTemperature2024.rds")
```

```{r}
write_csv(SingleAspiratedAirTemp$SAAT_30min, "SingleAirTemp.csv")
write_csv(TripleAspiratedAirTemp$TAAT_30min, "TripleAirTemp.csv")
write_csv(SoilRadiation$SRDDP_30min, "SoilRadiation.csv")
write_csv(SoilTemperature2022$ST_30_minute, "SoilTemperature2022*.csv")
write_csv(SoilTemperature2023$ST_30_minute, "SoilTemperature2023*.csv")
write_csv(SoilTemperature2024$ST_30_minute, "SoilTemperature2024*.csv")
```

## START

```{r}
SingleAirTemp <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SingleAirTemp.csv")
TripleAirTemp <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/TripleAirTemp.csv")
SoilRadiation <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SoilRadiation.csv")
RelativeHumidity <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQRHumid.csv")
AtmosphereIsotope <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQAtmos.csv")
SoilTemperature2022 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SoilTemperature2022*.csv")
SoilTemperature2023 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SoilTemperature2023*.csv")
SoilTemperature2024 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SoilTemperature2024*.csv")
```

```{r}
create_period_summary <- function(data,
                                  value_column,
                                  period_days) {
  start_date <- min(data$Date, na.rm = TRUE)
  data %>%
    mutate(
      PeriodGroup =
        floor(as.numeric(Date - start_date) / period_days)
    ) %>%
    group_by(PeriodGroup) %>%
    summarize(
      StartDate = min(Date),
      EndDate = max(Date),
      Mean = mean(.data[[value_column]], na.rm = TRUE),
      SD = sd(.data[[value_column]], na.rm = TRUE),
      N = sum(!is.na(.data[[value_column]])),
      .groups = "drop"
    )
}
```

## Soil Radiation

### Identifying Sunset v. Sunrise

```{r}
SoilRadiation <- SoilRadiation %>% 
  mutate(
    Daylight = difRadMean > 10
  )

Daylight <- SoilRadiation %>% 
  select(Daylight, endDateTime, startDateTime)


```

```{r}
# #making date column
# SoilRadiation <- SoilRadiation %>%
#   mutate(
#     Date = as.Date(endDateTime)
#   )
#   
# #making date into time
# SoilRadiation <- SoilRadiation %>% 
#   mutate(
#     Time = format(endDateTime, "%H:%M:%S")
#   )
```

```{r}
# Sunrise_Sunset <- SoilRadiation %>%
#   mutate(
#     daylight = dirRadMean > 10
#   ) %>%
#   filter(daylight) %>%
#   group_by(Date) %>%
#   summarise(
#     Sunrise = first(Time),
#     Sunset = last(Time),
#     .groups = "drop"
#   )
```

## Soil Temperature

### Incorporating daytime into soil temp measurements

```{r}
SoilTemp <- bind_rows(SoilTemperature2022, SoilTemperature2023, SoilTemperature2024)
```

```{r}
SoilAvgTemp <- SoilTemp %>% 
  group_by(startDateTime) %>%
  summarise(
    MeanSoilTemp = mean(soilTempMean, na.rm = TRUE),
    NumSensors = sum(!is.na(soilTempMean)),
    .groups = "drop"
  )
```

```{r}
SoilAvgTemp <- SoilAvgTemp %>% 
  left_join(
    Daylight %>% select(startDateTime, Daylight),
    by = "startDateTime"
  )

```

```{r}
SoilDaytimeAvg <- SoilAvgTemp %>% 
  mutate(
    DateTime = ymd_hms(startDateTime),
    Date = as.Date(DateTime)) %>% 
  filter(Daylight == TRUE)
```

```{r}
SoilTemp_3day <- create_period_summary(
  data = SoilDaytimeAvg,
  value_column = "MeanSoilTemp",
  period_days = 3
)

SoilTemp_7day <- create_period_summary(
  SoilDaytimeAvg,
  "MeanSoilTemp",
  7
)

SoilTemp_14day <- create_period_summary(
  SoilDaytimeAvg,
  "MeanSoilTemp",
  14
)
```

## Atmosphere temperature

### calculating averages

```{r}
SingleAvgAirTemp <- SingleAirTemp %>% 
  group_by(startDateTime) %>%
  summarise(
    SingleMeanAirTemp = mean(tempSingleMean, na.rm = TRUE),
    .groups = "drop"
  )

TripleAvgAirTemp <- TripleAirTemp %>% 
  group_by(startDateTime) %>%
  summarise(
    TripleMeanAirTemp = mean(tempTripleMean, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
CombinedAvgAirTemp <- SingleAvgAirTemp %>% 
  left_join(
    TripleAvgAirTemp %>% select(startDateTime, TripleMeanAirTemp),
    by = "startDateTime"
  )
```

```{r}
CombinedAvgAirTemp <- CombinedAvgAirTemp %>% 
  left_join(
    Daylight %>% select(startDateTime, Daylight),
    by = "startDateTime"
  )
```

```{r}
CombinedAvgAirTemp <- CombinedAvgAirTemp %>% 
  mutate(
    AvgAirTemp = ((SingleMeanAirTemp + TripleMeanAirTemp)/2)
  )
```

```{r}
CombinedAvgAirTemp <- CombinedAvgAirTemp %>% 
  mutate(
    DateTime = ymd_hms(startDateTime),
    Date = as.Date(DateTime)) %>% 
  filter(Daylight == TRUE)
```

```{r}
AirTemp_3day <- create_period_summary(
  data = CombinedAvgAirTemp,
  value_column = "AvgAirTemp",
  period_days = 3
)

AirTemp_7day <- create_period_summary(
  data = CombinedAvgAirTemp,
  value_column = "AvgAirTemp",
  period_days = 7
)

AirTemp_14day <- create_period_summary(
  data = CombinedAvgAirTemp,
  value_column = "AvgAirTemp",
  period_days = 14
)
```

## Humidity

```{r}
RelativeHumidity <- RelativeHumidity %>% 
  group_by(startDateTime) %>%
  summarise(
    RHMean = mean(RHMean, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
RelativeHumidity <- RelativeHumidity %>% 
  left_join(
    Daylight %>% select(startDateTime, Daylight),
    by = "startDateTime"
  )
```

```{r}
RelativeHumidity <- RelativeHumidity %>% 
  mutate(
    DateTime = ymd_hms(startDateTime),
    Date = as.Date(DateTime)) %>% 
  filter(Daylight == TRUE)
```

```{r}
RH_3day <- create_period_summary(
  data = RelativeHumidity,
  value_column = "RHMean",
  period_days = 3
)

RH_7day <- create_period_summary(
  data = RelativeHumidity,
  value_column = "RHMean",
  period_days = 7
)

RH_14day <- create_period_summary(
  data = RelativeHumidity,
  value_column = "RHMean",
  period_days = 14
)
```

## Atmospheric Isotopes

### d18O

```{r}
AtmosphereIsotope <- AtmosphereIsotope %>%
  rename(
    "startDateTime" = "timeBgn",
    "endDateTime" = "timeEnd"
  )
```

```{r}
d18OAtmosIsotopes <- AtmosphereIsotope %>% 
  group_by(startDateTime) %>% 
  summarise(
    Meand18OAtmosIsotopes = mean(data.isoH2o.dlta18OH2o.mean, na.rm = TRUE)
  )
```

```{r}
d18OAtmosIsotopes <- d18OAtmosIsotopes %>% 
  left_join(
    Daylight %>% select(startDateTime, Daylight),
    by = "startDateTime"
  )
```

```{r}
d18OAtmosIsotopes <- d18OAtmosIsotopes %>% 
  mutate(
    DateTime = ymd_hms(startDateTime),
    Date = as.Date(DateTime)) %>% 
  filter(Daylight == TRUE)
```

```{r}
dAd18O_3day <- create_period_summary(
  data = d18OAtmosIsotopes,
  value_column = "Meand18OAtmosIsotopes",
  period_days = 3
)

dAd18O_7day <- create_period_summary(
  data = d18OAtmosIsotopes,
  value_column = "Meand18OAtmosIsotopes",
  period_days = 7
)

dAd18O_14day <- create_period_summary(
  data = d18OAtmosIsotopes,
  value_column = "Meand18OAtmosIsotopes",
  period_days = 14
)
```

### d2H

```{r}
d2HAtmosIsotopes <- AtmosphereIsotope %>% 
  group_by(startDateTime) %>% 
  summarise(
    Meand2HAtmosIsotopes = mean(data.isoH2o.dlta2HH2o.mean, na.rm = TRUE)
  )
```

```{r}
d2HAtmosIsotopes <- d2HAtmosIsotopes %>% 
  left_join(
    Daylight %>% select(startDateTime, Daylight),
    by = "startDateTime"
  )
```

```{r}
d2HAtmosIsotopes <- d2HAtmosIsotopes %>% 
  mutate(
    DateTime = ymd_hms(startDateTime),
    Date = as.Date(DateTime)) %>% 
  filter(Daylight == TRUE)
```

```{r}
dAd2H_3day <- create_period_summary(
  data = d2HAtmosIsotopes,
  value_column = "Meand2HAtmosIsotopes",
  period_days = 3
)

dAd2H_7day <- create_period_summary(
  data = d2HAtmosIsotopes,
  value_column = "Meand2HAtmosIsotopes",
  period_days = 7
)

dAd2H_14day <- create_period_summary(
  data = d2HAtmosIsotopes,
  value_column = "Meand2HAtmosIsotopes",
  period_days = 14
)
```

### dS and dT

```{r}
ONAQdSdTPath <- "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ExampleCode/katarenamatos-NEONET/ONAQ_dS-and-dT.xlsx"

sheets <- excel_sheets(ONAQdSdTPath)

for(sheet in sheets) {
  
  ONAQdSdTSheet <- read_excel(ONAQdSdTPath, sheet = sheet)
  csv_name <- paste0(sheet, ".csv")
  write.csv(ONAQdSdTSheet, csv_name, row.names = FALSE)
}
```

```{r}
ONAQdSdT <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQ.csv")
```

```{r}
ONAQdSdT <- ONAQdSdT %>% 
  rename(
    "Sample_ID" = "Sample ID", 
    "Site_ID" = "Site ID",
    "d2HdS" = "dS2H Summary",
    "d18OdS" = "dS18O Summary",
    "d2HdT" = "dT2H Summary",
    "d18OdT" = "dT18O Summary",
    "SampleDate" = "Sample Date",
    "dS18O_SE" = "...6",
    "dS2H_SE" = "...9",
    "dS18O_Num" = "...7",
    "dS2H_Num" = "...10",
    "dT18O_SE" = "...12",
    "dT18O_Num" = "...13",
    "dT2H_SE" = "...15",
    "dT2H_Num" = "...16"
  )
```

```{r}
ONAQdS <- ONAQdSdT %>% 
  select(Sample_ID, Site_ID, SampleDate, d18OdS, dS18O_SE, dS18O_Num, d2HdS, dS2H_SE, dS2H_Num)

ONAQdT <- ONAQdSdT %>% 
  select(Sample_ID, Site_ID, SampleDate, d18OdT, dT18O_SE, dT18O_Num, d2HdT, dT2H_SE, dT2H_Num)
```

```{r}
dS2Hplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d2HdS)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dS2Hplt

dT2Hplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d2HdT)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dT2Hplt

dSd18Oplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d18OdS)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dSd18Oplt

dTd18Oplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d18OdT)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dTd18Oplt
```

## Monte Carlo Analysis

```{r}
# mean_SoilTemp <- SoilTemp_3day$Mean [i]
# sd_SoilTemp <- SoilTemp_3day$SD [i]
# 
# mean_AirTemp <- AirTemp_3day$Mean [i]
# sd_AirTemp <- AirTemp_3day$SD [i]
# 
# mean_RH <- RH_3day$Mean [1]
# sd_RH <- RH_3day$SD [1]
# 
# mean_dA18O <- dAd18O_3day$Mean [1]
# sd_dAd18O <- dAd18O_3day$SD [1]
# 
# mean_dAd2H <- dAd2H_3day$Mean [1]
# sd_dAd2H <- dAd2H_3day$Mean [1]
```

```{r}
#soil temp monte carlo simulation
MC_SoilTemp <- vector(
  mode = "list",
  length = nrow(SoilTemp_3day)
)
for(i in 1:nrow(SoilTemp_3day)){
  MC_SoilTemp[[i]] <- rnorm(
    n = 1000,
    mean = SoilTemp_3day$Mean[i],
    sd = SoilTemp_3day$SD[i]
  )}

#air temp
MC_AirTemp <- vector(
  mode = "list",
  length = nrow(AirTemp_3day)
)
for(i in 1:nrow(AirTemp_3day)){
  
  MC_AirTemp[[i]] <- rnorm(
    n = 1000,
    mean = AirTemp_3day$Mean[i],
    sd = AirTemp_3day$SD[i]
  )}

#relative humidity
MC_RH <- vector(
  mode = "list",
  length = nrow(RH_3day)
)
for(i in 1:nrow(RH_3day)){
  MC_RH[[i]] <- rnorm(
    n = 1000,
    mean = RH_3day$Mean[i],
    sd = RH_3day$SD[i]
  )}

#atmospheric isotopes oxygen-18
MC_dAd18O <- vector(
  mode = "list",
  length = nrow(dAd18O_3day)
)
for(i in 1:nrow(dAd18O_3day)){
  MC_dAd18O[[i]] <- rnorm(
    n = 1000,
    mean = dAd18O_3day$Mean[i],
    sd = dAd18O_3day$SD[i]
  )}

#atmospheric isotopes hydrogen-2
MC_dAd2H <- vector(
  mode = "list",
  length = nrow(dAd2H_3day)
)
for(i in 1:nrow(dAd2H_3day)){
  MC_dAd2H[[i]] <- rnorm(
    n = 1000,
    mean = dAd2H_3day$Mean[i],
    sd = dAd2H_3day$SD[i]
  )}

#soil water isotopes oxygen-18
MC_dSd18O <- vector(
  mode = "list",
  length = nrow(ONAQdS)
)
for(j in 1:nrow(ONAQdS)){
  MC_dSd18O[[j]] <- rnorm(
    n = 1000,
    mean = ONAQdS$d18OdS,
    sd = ONAQdS$dS18O_SE
  )
}

#soil water isotopes hydrogen-2
MC_dSd2H <- vector(
  mode = "list",
  length = nrow(ONAQdS)
)
for(j in 1:nrow(ONAQdS)){
  MC_dSd2H[[j]] <- rnorm(
    n = 1000,
    mean = ONAQdS$d2HdS,
    sd = ONAQdS$dS2H_SE
  )
}

#plant water isotopes oxygen-18
MC_dTd18O <- vector(
  mode = "list",
  length = nrow(ONAQdT)
)
for(j in 1:nrow(ONAQdT)){
  MC_dTd18O[[j]] <- rnorm(
    n = 1000,
    mean = ONAQdT$d18OdT,
    sd = ONAQdT$dT18O_SE
  )
}

#plant water isotopes hydrogen-2
MC_dTd2H <- vector(
  mode = "list",
  length = nrow(ONAQdT)
)
for(j in 1:nrow(ONAQdT)){
  MC_dTd2H[[j]] <- rnorm(
    n = 1000,
    mean = ONAQdT$d2HdT,
    sd = ONAQdT$dT2H_SE
  )
}
```

### Craig Gordon

#### alpha star

##### delta 18O

$$\alpha{^{+18O}}=\frac{1}{\exp\left(\frac{-7.685+ 6.7123\left(\frac{10^3}{T+273.15}\right)- 1.6664\left(\frac{10^6}{(T+273.15)^2}\right)+ 0.35041\left(\frac{10^9}{(T+273.15)^3}\right)}{10^3}\right)}$$

```{r}

MC_alphastar_d18O <- vector(
  mode = "list",
  length = nrow(SoilTemp_3day)
)

for(i in 1:nrow(SoilTemp_3day)) {

 MC_alphastar_d18O[[i]] <- 1/(exp(
  (
    -7.685 +
    6.7123*(10^3/(MC_SoilTemp[[i]] + 273.15)) -
    1.6664*(10^6/(MC_SoilTemp[[i]] + 273.15)^2) +
    0.35041*(10^9/(MC_SoilTemp[[i]] + 273.15)^3)
  ) / 10^3
))}
```

##### delta 2H

$$\alpha{^{+2H}}=\frac{1}{\exp\left(\frac{1158.8\left(\frac{(T+273.15)^3}{10^9}\right)- 1620.1\left(\frac{(T+273.15)^2}{10^6}\right)+ 794.84\left(\frac{T+273.15}{10^3}\right)- 161.04+ 2.9992\left(\frac{10^9}{(T+273.15)^3}\right)}{10^3}\right)}$$

```{r}
MC_alphastar_d2H <- vector(
  mode = "list",
  length = nrow(SoilTemp_3day)
)

for(i in 1:nrow(SoilTemp_3day)) {

 MC_alphastar_d2H[[i]] <- 1/exp(
  (
    1158.8 * ((MC_SoilTemp[[i]] + 273.15)^3 / 10^9) -
    1620.1 * ((MC_SoilTemp[[i]] + 273.15)^2 / 10^6) +
    794.84 * ((MC_SoilTemp[[i]] + 273.15) / 10^3) -
    161.04 +
    2.9992 * (10^9 / (MC_SoilTemp[[i]] + 273.15)^3)
  ) / 10^3
)}
```

#### epsilon star - equilibrium isotopic separations

##### d18O

$$
\varepsilon^{+18O} = (\alpha^{+18O}-1)*10^{3}‰
$$

```{r}
MC_epsilonstar_d18O <- vector(
  mode = "list",
  length = nrow(MC_alphastar_d18O)
)

for(i in 1:nrows(MC_alphastar_d18O)) {
  MC_epsilonstar_d18O[[i]] <- ((MC_alphastar_d18O - 1)*10^3)
}
```

##### d2H

$$
\varepsilon^{+2H} = (\alpha^{+2H}-1)*10^{3}‰
$$

```{r}
MC_epsilonstar_d2H <- vector(
  mode = "list",
  length = nrows(MC_alphastar_d2H)
)

for(i in 1:nrows(MC_alphastar_d2H)) {
  MC_epsilonstar_d2H[[i]] <- ((MC_alphastar_d2H - 1)*10^3)
}
```

#### humidity

converts the measured humidity in the air to what it likely is at the soil surface based on the temperature of each and therefore how much water the system can hold

-   if the soil is hotter than the environment than it is less humid

-   for Utah lake, zanazzi uses atmospheric humidity and does not convert based on temperature at lake surface

##### air vapor pressure

$$e_{sat,air}= \frac{6.11*10^{\frac{7.5T_a}{237.3+T_a}}}{10}$$

```{r}
MC_airvappres <- vector(
  mode = "list",
  length = nrows(MC_RH)
)
for(i in nrows(MC_RH)) {
  MC_airvappres[[i]] <- (6.11*10^((7.5*MC_AirTemp[[i]])/(237.3+MC_AirTemp[[i]])))/10 
}
```

##### soil vapor pressure

$$e_{sat,soil} = \frac{6.11*10^{\frac{7.5T_s}{237.3+T_s}}}{10} $$

```{r}
MC_soilvappres <- vector(
  mode = "list",
  length = nrows(MC_SoilTemp)
)
for(i in nrows(MC_SoilTemp)) {
  MC_soilvappres[[i]] <- (6.11*10^((7.5*MC_SoilTemp[[i]])/(237.3+MC_SoilTemp[[i]])))/10 
}
```

##### humidity fraction

$$humidity = \frac{RH}{100}*\frac{e_{sat,air}}{e_{sat,soil}}$$

```{r}
MC_hA <- vector(
  mode = "list",
  length = nrows(MC_RH)
)
for(i in nrows(MC_RH)){
  MC_hA[[i]] <- ((MC_RH[[i]])/100)*((MC_airvappres[[i]]/MC_soilvappres[[i]]))
}
```

#### sigma k - kinetic isotopic fractionation factors

d18O

$$\varepsilon_K^{18O} = \theta*C_K^{18O}*(1-h)‰$$

##### 

```{r}
MC_epsilonK_18O <-  vector(
  mode= "list",
  length = nrows(MC_hA)
)
for (i in 1:nrows(MC_hA)) {
  MC_epsilonK[[i]] <- 1*14.2*(1-MC_hA[[i]])
  
}
```

##### d2H

$$\varepsilon_K^{18O} = \theta*C_K^{2H}*(1-h)‰$$

```{r}
MC_epsilonK_2H <-  vector(
  mode= "list",
  length = nrows(MC_hA)
)
for (i in 1:nrows(MC_hA)) {
  MC_epsilonK[[i]] <- 1*12.5*(1-MC_hA[[i]])
  
}
```

#### Craig Gordon equation

##### d18O

$$\delta_E^{18O} = \frac{(\alpha_{18O}*\delta_S^{18O}-h_A\delta_A^{18O}) - \frac{\varepsilon_{+}^{18O} + \varepsilon_k^{18O}}{1000}}{(1-h_A)+\frac{\varepsilon_k^{18O}}{1000}}$$

```{r}
MC_CG <- vector(
  mode = "list",
  length = nrows(MC_dSd18O)
)

for(i in nrows(MC_dSd18O)){
  MC_CG[[i]] <- ((MC_alphastar_d18O*MC_dSd18O-MC_hA*MC_dAd18O)-((MC_epsi))
}
```
