---
title: "MatosAnalysis"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(latex2exp)
library(readxl)
library(egg)
library(plotly)
library(readr)
library(neonUtilities) #download and manage NEON data (in-house)
library(neonOS) #NEON OS data
library(terra) #general geological data
library(ggplot2)
library(colorspace)
library(viridis)
library(sf)
library(tigris)
library(leaflet)
library(maps)
library(crosstalk)
library(lubridate)
```

## Loading New Data Products

Need:

-   atmos vapor concentration

-   soil temperature

-   daily air temp

-   solar radiation

Have:

-   atmos isotope concentration

-   precipitation amount

-   precipitation isotope concentration

-   relative humidity

```{r}
# soil radiation
SoilRadiation <- loadByProduct(
  dpID = "DP1.00014.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE)
```

```{r}
#single aspirated air temp
SingleAspiratedAirTemp <- loadByProduct(
  dpID = "DP1.00002.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE)
```

```{r}
#triple aspirated air temp
TripleAspiratedAirTemp <- loadByProduct(
  dpID = "DP1.00003.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE)
```

```{r}
#soil temp
SoilTemperature2022 <- loadByProduct(
  dpID = "DP1.00041.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2022-12",
  package = "basic",
  check.size = FALSE
)
  
```

```{r}
SoilTemperature2023 <- loadByProduct(
  dpID = "DP1.00041.001",
  site = "ONAQ",
  startdate = "2023-01",
  enddate = "2023-06",
  package = "basic",
  check.size = FALSE
)
```

```{r}
write_rds(SoilTemperature2022$ST_30_minute, "SoilTemperature2022.rds")
write_rds(SoilTemperature2023$ST_30_minute, "SoilTemperature2023.rds")
write_rds(SoilTemperature2024$ST_30_minute, "SoilTemperature2024.rds")
```

```{r}
write_csv(SoilTemperature2022$ST_30_minute, "SoilTemperature2022.csv")
write_csv(SoilTemperature2023$ST_30_minute, "SoilTemperature2023.csv")
write_csv(SoilTemperature2024$ST_30_minute, "SoilTemperature2024.csv")
```

```{r}
SoilTemperature2024 <- loadByProduct(
  dpID = "DP1.00041.001",
  site = "ONAQ",
  startdate = "2024-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE
)
```

```{r}
write_csv(SingleAspiratedAirTemp$SAAT_30min, "SingleAirTemp.csv")
write_csv(TripleAspiratedAirTemp$TAAT_30min, "TripleAirTemp.csv")
write_csv(SoilRadiation$SRDDP_30min, "SoilRadiation.csv")
```

## START

```{r}
SingleAirTemp <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SingleAirTemp.csv")
TripleAirTemp <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/TripleAirTemp.csv")
SoilRadiation <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SoilRadiation.csv")
RelativeHumidity <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQRHumid.csv")
AtmosphereIsotope <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQAtmos.csv")
SoilTemperature2022 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SoilTemperature2022.csv")
SoilTemperature2023 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SoilTemperature2023.csv")
SoilTemperature2024 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SoilTemperature2024.csv")
```

## Soil Radiation

### Identifying Sunset v. Sunrise

```{r}
SoilRadiation <- SoilRadiation %>% 
  mutate(
    Daylight = difRadMean > 10
  )

Daylight <- SoilRadiation %>% 
  select(Daylight, endDateTime)


```

```{r}
#making date column
SoilRadiation <- SoilRadiation %>%
  mutate(
    Date = as.Date(endDateTime)
  )
  
#making date into time
SoilRadiation <- SoilRadiation %>% 
  mutate(
    Time = format(endDateTime, "%H:%M:%S")
  )
```

```{r}
Sunrise_Sunset <- SoilRadiation %>%
  mutate(
    daylight = dirRadMean > 10
  ) %>%
  filter(daylight) %>%
  group_by(Date) %>%
  summarise(
    Sunrise = first(Time),
    Sunset = last(Time),
    .groups = "drop"
  )
```

## Soil Temperature

### Incorporating daytime into soil temp measurements

```{r}
SoilTemp <- bind_rows(SoilTemperature2022, SoilTemperature2023, SoilTemperature2024)
```

```{r}
SoilAvgTemp <- SoilTemp %>% 
  group_by(endDateTime) %>%
  summarise(
    MeanSoilTemp = mean(soilTempMean, na.rm = TRUE),
    NumSensors = sum(!is.na(soilTempMean)),
    .groups = "drop"
  )
```

```{r}
SoilAvgTemp <- SoilAvgTemp %>% 
  left_join(
    Daylight %>% select(endDateTime, Daylight),
    by = "endDateTime"
  )

```

```{r}
SoilDaytimeAvg <- SoilAvgTemp %>% 
  mutate(
    DateTime = ymd_hms(endDateTime),
    Date = as.Date(DateTime)) %>% 
  filter(Daylight == TRUE)
```

```{r}
create_period_summary <- function(data,
                                  value_column,
                                  period_days) {

  start_date <- min(data$Date, na.rm = TRUE)

  data %>%
    mutate(
      PeriodGroup =
        floor(as.numeric(Date - start_date) / period_days)
    ) %>%
    group_by(PeriodGroup) %>%
    summarize(
      StartDate = min(Date),
      EndDate = max(Date),

      Mean = mean(.data[[value_column]], na.rm = TRUE),

      SD = sd(.data[[value_column]], na.rm = TRUE),

      N = sum(!is.na(.data[[value_column]])),

      .groups = "drop"
    )
}
```

```{r}
Ts_3day <- create_period_summary(
  data = SoilDaytimeAvg,
  value_column = "MeanSoilTemp",
  period_days = 3
)

SoilTemp_7day <- create_period_summary(
  SoilDaytimeAvg,
  "MeanSoilTemp",
  7
)

SoilTemp_14day <- create_period_summary(
  SoilDaytimeAvg,
  "MeanSoilTemp",
  14
)
```

## Atmosphere temperature

### calculating averages

```{r}
SingleAvgAirTemp <- SingleAirTemp %>% 
  group_by(endDateTime) %>%
  summarise(
    SingleMeanAirTemp = mean(tempSingleMean, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
CombinedAvgAirTemp <- SingleAvgAirTemp %>% 
  left_join(
    TripleAirTemp %>% select(endDateTime, tempTripleMean),
    by = "endDateTime"
  )
```

```{r}
CombinedAvgAirTemp <- CombinedAvgAirTemp %>% 
  left_join(
    Daylight %>% select(endDateTime, Daylight),
    by = "endDateTime"
  )
```

```{r}
CombinedAvgAirTemp <- CombinedAvgAirTemp %>% 
  mutate(
    AvgAirTemp = ((SingleMeanAirTemp + tempTripleMean)/2)
  )
```

```{r}
CombinedAvgAirTemp <- CombinedAvgAirTemp %>% 
  mutate(
    DateTime = ymd_hms(endDateTime),
    Date = as.Date(DateTime)) %>% 
  filter(Daylight == TRUE)
```

```{r}
Ta_3day <- create_period_summary(
  data = CombinedAvgAirTemp,
  value_column = "AvgAirTemp",
  period_days = 3
)

Ta_7day <- create_period_summary(
  data = CombinedAvgAirTemp,
  value_column = "AvgAirTemp",
  period_days = 7
)

Ta_14day <- create_period_summary(
  data = CombinedAvgAirTemp,
  value_column = "AvgAirTemp",
  period_days = 14
)
```

## Humidity

```{r}
RelativeHumidity <- RelativeHumidity %>% 
  group_by(endDateTime) %>%
  summarise(
    RHMean = mean(RHMean, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
RelativeHumidity <- RelativeHumidity %>% 
  left_join(
    Daylight %>% select(endDateTime, Daylight),
    by = "endDateTime"
  )
```

```{r}
RelativeHumidity <- RelativeHumidity %>% 
  mutate(
    DateTime = ymd_hms(endDateTime),
    Date = as.Date(DateTime)) %>% 
  filter(Daylight == TRUE)
```

```{r}
RH_3day <- create_period_summary(
  data = CombinedAvgAirTemp,
  value_column = "AvgAirTemp",
  period_days = 3
)

RH_7day <- create_period_summary(
  data = CombinedAvgAirTemp,
  value_column = "AvgAirTemp",
  period_days = 7
)

RH_14day <- create_period_summary(
  data = CombinedAvgAirTemp,
  value_column = "AvgAirTemp",
  period_days = 14
)
```

## Atmospheric Isotopes

### d18O

```{r}
AtmosphereIsotope <- AtmosphereIsotope %>% 
  group_by(timeEnd) %>% 
  summarise(
    Meand18OAtmosIsotopes = mean(data.isoH2o.dlta18OH2o.mean, na.rm = TRUE)
  )

```

```{r}
Avgd18OAtmosIsotopes <- Avgd18OAtmosIsotopes %>% 
  left_join(
    Daylight %>% select(endDateTime, Daylight),
    by = "timeEnd"
  )
```

### d2H

```{r}
Avgd2HAtmosIsotopes <- AtmosphereIsotope %>% 
  group_by(timeEnd) %>% 
  summarise(
    MeanAtmosIsotopes = mean(data.isoH2o.dlta2HH2o.mean, na.rm = TRUE)
  )
```

```{r}
Avgd2HAtmosIsotopes <- Avgd2HAtmosIsotopes %>% 
  mutate(
    YearMonth = format(timeEnd, "%Y-%m")
  )
```

```{r}
Avgd2HAtmosIsotopes <- Avgd2HAtmosIsotopes %>% 
  left_join(
    Daylight %>% select(endDateTime, Daylight),
    by = "endDateTime"
  )
```

### dS and dT

```{r}
ONAQdSdTPath <- "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ExampleCode/katarenamatos-NEONET/ONAQ_dS-and-dT.xlsx"

sheets <- excel_sheets(ONAQdSdTPath)

for(sheet in sheets) {
  
  ONAQdSdTSheet <- read_excel(ONAQdSdTPath, sheet = sheet)
  csv_name <- paste0(sheet, ".csv")
  write.csv(ONAQdSdTSheet, csv_name, row.names = FALSE)
}
```

```{r}
ONAQdSdT <- read_csv("ONAQ.csv")
```

```{r}
ONAQdSdT <- ONAQdSdT %>% 
  rename(
    "d2HdS" = "dS2H Summary",
    "d18OdS" = "dS18O Summary",
    "d2HdT" = "dT2H Summary",
    "d18OdT" = "dT18O Summary",
    "SampleDate" = "Sample Date"
  )
```

```{r}
dS2Hplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d2HdS)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dS2Hplt

dT2Hplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d2HdT)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dT2Hplt

dSd18Oplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d18OdS)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dSd18Oplt

dTd18Oplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d18OdT)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dTd18Oplt
```
