---
title: "UROPSummer2026"
format: html
editor: visual
---

```{r}
#did this so it would use files from UROP Summer 2026 but doesn't work and files to still automatically go to the git hub file, therefore need to specify when calling the file in individual code lines
setwd("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026")
```

```{r}
library(tidyverse)
library(latex2exp)
library(readxl)
library(egg)
library(plotly)
library(readr)
library(neonUtilities) #download and manage NEON data (in-house)
library(neonOS) #NEON OS data
library(terra) #general geological data
library(ggplot2)
library(colorspace)#got for using more colors, don't really use
library(viridis)#virids color palette
library(sf)#need for interactive map
library(tigris)#interactive map
library(leaflet)#need for satellite image of interactive map
library(maps)#interactive map
library(crosstalk)
```

## Test Run - April 2024 Only

```{r}
#april 2024 atmosphere data, test run so could deal with smaller data set. this generates the #h5 files on the computer but need to be picked apart manually on the computer manually
zipsByProduct(
  dpID = "DP4.00200.001",
  site = "ONAQ",
  startdate = "2024-04",
  enddate = "2024-04",
  package = "basic",
  check.size = FALSE
)
```

```{r}
setwd("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP")
getwd()

```

```{r}
#stacks the files to be a df in R from the manually unstacked h5 files. file path leads to the h5 files that i manually placed into this folder
AprilEddy <- stackEddy(
  filepath = "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/filesToStack00200/AprilSampleh5",
  level = "dp01",
  var = "tempAirTop",
  avg = 30
)

```

```{r}
#creates april 2024 only atmosphere isotopes csv file
write.csv(Eddy$ONAQ, file = "ONAQAprilAtm.csv")

```

```{r}
#reads april 2024 only atmosphere isotopes csv file
ONAQTRUEAprilAtm <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/ONAQAprilAtm.csv")
```

## All

### Atm and Air Temp

```{r}
#same as april 2024 above. pulls out zip file of all eddy covalence data. need to go in #manually and spereate out everything into its own file with all the h5 files.
zipsByProduct(
  dpID = "DP4.00200.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE
)
```

```{r}
#pulls out atmosphere isotopes from eddy covalence data and puts it into its own df inside R
AtmosEddy <- stackEddy(
  filepath = "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQAtmosH5",
  level = "dp01",
  var = "isoH2o",
  avg = 30
)

#does the same thing as the atmosphere, pulls out air temp from eddy covalence data and then organizes into its own df
AirTempEddy <- stackEddy(
  filepath = "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQAtmosH5",
  level = "dp01",
  var = "tempAirTop",
  avg = 30
)
```

```{r}
#writes the csv file for both atmosphere isotopes data and air temp data from eddy covalence from the df made above from the stack eddy func. these now exsit on the computer
write_csv(AtmosEddy$ONAQ, file = "ONAQAtmos.csv")
write_csv(AirTempEddy$ONAQ, file = "ONAQAirTemp.csv")
```

### Soil Water Iso

```{r}
#pulls data from the soil data from NEON. did the same thing where it stacked together all of the zip files and had to back and manually pull them out. but this isn't the eddy covalence, its for the soil data on neon
zipsByProduct(
  dpID = "DP1.10086.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE
  )
```

```{r}
#pulls soil isotope data from the h5 files and creates new df
SoilIsoEddy <- stackEddy(
  filepath = "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/filesToStack10086",
  level = "dp01",
  var = "soil",
  avg = 30
)
```

Need to add in csv file for this \*maybe

couldn't find the actual soil isotope data for this. probably don't need for utah lake calculations. might need for matos but they use their own df which is found in their example code file

### Humidity

```{r}
#load humidity df
ONAQhumidity <- loadByProduct(
  dpID = "DP1.00098.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE
  
)

#as tibble function creates own df with the 30 min measurments then the write csv creates the specific csv file in urop sum 2026 file
ONAQHumid <- as_tibble(ONAQhumidity$RH_30min)
write_csv(ONAQHumid, "ONAQRHumid.csv")
```

### Precipitation

```{r}
#same as humidity above, loads data from cloud/website into the r environment
ONAQprecipitation <- loadByProduct(dpID = "DP1.00038.001", 
                                   site = "ONAQ",
                                   startdate = "2022-01",
                                   enddate = "2024-12",
                                   check.size = FALSE)

#now creates df with just the precipitation isotope measurements, and then turns this into a csv file that will now exist on the computer
ONAQPrecip <- as_tibble(ONAQprecipitation$wdi_isoPerSample)
write_csv(ONAQPrecip, "ONAQPrecip.csv")
```

# Start

### Atmosphere

```{r}
#loads in csv file for onaq atmospheric isotopes (again have to include the whole string because of weird file path stuff), TRUE label is not edited at all and is straight original version
ONAQTRUEAtmos <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQAtmos.csv")
```

```{r}
#modifies names of columns for easier work later and ease of understanding the data
ONAQAtmos <- ONAQTRUEAtmos %>% 
  rename(
    "d18Omean" = "data.isoH2o.dlta18OH2o.mean",
    "d2Hmean" = "data.isoH2o.dlta2HH2o.mean",
    "StartTime" = "timeBgn",
    "EndTime" = "timeEnd",
    "d18Omin" = "data.isoH2o.dlta18OH2o.min",
    "d18Omax" = "data.isoH2o.dlta18OH2o.max",
    "d2Hmin" = "data.isoH2o.dlta2HH2o.min",
    "d2Hmax" = "data.isoH2o.dlta2HH2o.max",
    "d18OSampleNumber" = "data.isoH2o.dlta18OH2o.numSamp",
    "d2HSampleNumber" = "data.isoH2o.dlta2HH2o.numSamp",
    "PressureMean" = "data.isoH2o.pres.mean",
    "PressureMin" = "data.isoH2o.pres.min",
    "PressureMax" = "data.isoH2o.pres.max",
    "PressureSampleNum" = "data.isoH2o.pres.numSamp",
    "TempMean" = "data.isoH2o.temp.mean",
    "TempMin" = "data.isoH2o.temp.min",
    "TempMax" = "data.isoH2o.temp.max",
    "DryMoleRatioMean" = "data.isoH2o.rtioMoleDryH2o.mean",
    "DryMoleRatioMin" = "data.isoH2o.rtioMoleDryH2o.min",
    "DryMoleRatioMax" = "data.isoH2o.rtioMoleDryH2o.max",
    "WetMoleRatioMean" = "data.isoH2o.rtioMoleWetH2o.mean",
    "WetMoleRatioMin" = "data.isoH2o.rtioMoleWetH2o.min",
    "WetMoleRatioMax" = "data.isoH2o.rtioMoleWetH2o.max"
    
  )
```

```{r}
#getting rid of unnecssary columns, specifically issue logs or repeats of data columns that came in a different format
ONAQAtmos <- ONAQAtmos %>% 
  select(horizontalPosition, verticalPosition, StartTime, EndTime, d18Omean, d18Omin, d18Omax, d18OSampleNumber, d2Hmean, d2Hmin, d2Hmax, d2HSampleNumber, PressureMean, PressureMin, PressureMax, PressureSampleNum, TempMean, TempMax, TempMin, DryMoleRatioMean, DryMoleRatioMin, DryMoleRatioMax, WetMoleRatioMean, WetMoleRatioMin, WetMoleRatioMax)

#adding row numbers
ONAQAtmos <- ONAQAtmos %>% 
  mutate(
    Row = row_number()
  )

#adding date of collection (20XX-XX)
ONAQAtmos <- ONAQAtmos %>%
  mutate(
    Date = format(StartTime, "%Y-%m")
  )

#getting rid of NA values
#getting rid of anything NA for d18O sample number because this what i first found to be inconsistent
ONAQAtmos <- ONAQAtmos %>% 
  filter(d18OSampleNumber != 0)
#gets rid of any NA with d18O isotope measurments as extra precaution
ONAQAtmos <- ONAQAtmos %>% 
  filter(!is.na(d18Omean))
#gets rid of any NA with d2H isotope measurments as extra precaution
ONAQAtmos <- ONAQAtmos %>% 
  filter(!is.na(d2Hmean))
```

```{r}
#originally intended to calculate humidity here but instead chose to use the actual data set from NEON tracking humidity so decided to not use this caluclation and voided it from the code
#ONAQAtmos <- ONAQAtmos %>% 
  #mutate(
    #humidityratio = ((WetMoleRatioMean / DryMoleRatioMean) * 0.677)
  #)
```

### Checks

```{r}
#created a data frame checking the vertical position variable in the atmosphere df and what the different positions were and their effect on if the data was NA or not
#vertcheck <- ONAQAtmos %>% 
  #filter(verticalPosition == "h2oMed")

#vertcheck <- vertcheck %>% 
  #filter(!is.na(d18Omean))

```

### Precipitation

```{r}
#load in precipation csv, same as atmosphere where have to have long file name string, also same as above where true is unchanged and original file 
ONAQTRUEPrecip <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQPrecip.csv")

```

```{r}
#renamed columns to be easier to use and understand
ONAQPrecip <- ONAQTRUEPrecip %>% 
  rename(
    "d18O" = "d18OWater",
    "d2H" = "d2HWater",
    "d18Osd" = "d18OsdWater",
    "d2Hsd" = "d2HsdWater",
    "d2Huncert" = "d2HWaterUncert",
    "d18Ouncert" = "d18OWaterUncert"
  )

#adding row numbers
ONAQPrecip <- ONAQPrecip %>% 
  mutate(
    Row = row_number()
  )

#adding date (20XX-XX)
ONAQPrecip <- ONAQPrecip %>%
  mutate(
    Date = format(collectDate, "%Y-%m")
  )
```

```{r}
#getting rid of unwanted columns (stuff to do with issue log or other variables not needed)
ONAQPrecip <-  ONAQPrecip %>% 
  select(collectDate, d18O, d2H, d18Osd, d2Hsd, d18Ouncert, d2Huncert, Row, Date)
```

### Humidity

```{r}
#load in humidity, same as above need file string to find on computer and true is unedited version
ONAQTRUEHumid <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQRHumid.csv")
```

```{r}
#rename column names for easier use and understanding
ONAQHumid <- ONAQTRUEHumid %>% 
  rename(
    "StartTime" = "startDateTime",
    "EndTime" = "endDateTime",
    "RHMin" = "RHMinimum",
    "RHMax" = "RHMaximum",
    "MeanTempC" = "tempRHMean",
    "MinTempC" = "tempRHMinimum",
    "MaxTempC" = "tempRHMaximum",
    "dewTempMin" = "dewTempMinimum",
    "dewTempMax" = "dewTempMaximum"
  )


```

```{r}

#adding fahrenheit
ONAQHumid <- ONAQHumid %>% 
  mutate(
    Temp_F = MeanTempC * (9/5) + 32)

#adding kelvin
ONAQHumid <- ONAQHumid %>% 
  mutate(
    Temp_Kelvin = MeanTempC + 273.15)
#kelvin is used later for other calculations, and added fahrenheit to better understand the data

#adding row numbers
ONAQHumid <- ONAQHumid %>% 
  mutate(
    Row = row_number()
  )

#adding date (20XX-XX)
ONAQHumid <- ONAQHumid %>%
  mutate(
    Date = format(StartTime, "%Y-%m"))

#choosing wanted columns to keep in df, gettind rid of issue log or random data collected that isn't needed
ONAQHumid <- ONAQHumid %>% 
  select(
    StartTime, EndTime, horizontalPosition, verticalPosition, RHMean, RHMin, RHMax, RHVariance, RHNumPts, RHExpUncert, RHStdErMean, MeanTempC, MinTempC, MaxTempC, tempRHVariance, tempRHNumPts, tempRHExpUncert, tempRHStdErMean, dewTempMean, dewTempMin, dewTempMax, Row, Date, Temp_F, Temp_Kelvin
  )
```

```{r}

#data frame for 40 ft tower measurment of humidity
ONAQHumid40 <- ONAQHumid %>% 
  filter(verticalPosition == "040")

#data frame for 0ft tower measurment of humidity
ONAQHumid0 <- ONAQHumid %>% 
  filter(verticalPosition == "000")
```

### Utah Lake Isotopes

#### From WaterIsotopes.org

```{r}
#data from water isotopes.org, no date restrictions, no type restrictions, restricted by lat and long to only get utah county data - date range is 1990 - 2024
#wanted this data to get a picture of all data collected and then sample from this later for monte carlo
AllUtahCountyData <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/AllUtahCountyData.csv")
```

```{r}
#renaming some columns for ease and understanding
AllUtahCountyData <- AllUtahCountyData %>%
  rename(
  "Elevation" = "Elevation_mabsl",
  "SD2H" = "d2H_Analytical_SD",
  "SD18O" = "d18O_Analytical_SD"
)
```

```{r}
#getting rid of unnecessary columns that are unneccessary
AllUtahCountyData <- AllUtahCountyData %>% 
  select(-Start_Date, -Start_Time_Zone, -Collection_Time_Zone, -Phase, -Depth_meters, -Sample_Comments, -D17O, -D17O_Analytical_SD, -WI_Analysis_Source, -Site_Comments,  )
```

```{r}
#creating main df, getting rid of sample types named below only leaving lake, river, and spring samples
UtahLakeSamples <- AllUtahCountyData %>% 
  filter(!Type %in% c("Tap", "Milk", "Sprinkler", "Ground"))

```

```{r}
#adding date (20XX-XX)
UtahLakeSamples <- UtahLakeSamples %>%
  mutate(
    Date = format(Collection_Date, "%Y-%m"))

#adding Year column
UtahLakeSamples <- UtahLakeSamples %>% 
  mutate(
    Year = format(Collection_Date, "%Y")
  )

#creating DONOTUSE row
UtahLakeSamples <- UtahLakeSamples %>% 
  mutate(
    DONOTUSE = FALSE
  )

```

#### From Zanazzi

```{r}
#loading in the data from zanazzi et al. 2020 
UtahLakeZanazziPath <- "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/UtahLake2014-2016Zanazzi.xlsx"

#this df was weird and had lots of sheets inside the actaul file, so had to seperate out all the sheets inorder to use the one i wanted
#names object sheet that lists all of the sheets in the file
sheets <- excel_sheets(UtahLakeZanazziPath)

#creates individual files for each sheet in the file
for(sheet in sheets) {
  
  UtahLakeZanazziSheet <- read_excel(UtahLakeZanazziPath, sheet = sheet)
  csv_name <- paste0(sheet, ".csv")
  write.csv(UtahLakeZanazziSheet, csv_name, row.names = FALSE)
}


```

```{r}
#reads each of these files of individual sheets, but only need sheet 1 because these are all repeats individual collections
UtahLakeZanazziS2 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/Table S2.csv")

UtahLakeZanazziS3 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/Table S3.csv")

UtahLakeZanazziS4 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/Table S4.csv")

UtahLakeZanazziS5 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/Table S5.csv")
```

```{r}
#Zanazzi et al 2020 data, combination of all collected data so only read this csv file
UtahLakeZanazziS1 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/Table S1.csv")
```

```{r}

#rename func to make easier and consistent columns
UtahLakeZanazziS1 <- UtahLakeZanazziS1 %>% 
  rename(
    "Site_Name" = "...2",
    "Latitude" = "...3",
    "Longitude" = "...4",
    "Collection_Date" = "...5",
    "d2H" = "...6",
    "SD2H" = "...7",
    "d18O" = "...8",
    "SD18O" = "...9"
  )

#if d2h is NA, don't include
UtahLakeZanazziS1 <- UtahLakeZanazziS1 %>% 
  filter(!is.na(d2H)) 
#gets rid of first row because it was either blank or had column labels
UtahLakeZanazziS1 <- UtahLakeZanazziS1[-c(1), ]

#got rid of these columns because they were werid format issues
UtahLakeZanazziS1 <- UtahLakeZanazziS1 %>% select(-`Table S1. Biweekly Samples`)
```

```{r}
#turns collection data into a numeric value becasue originally was character 
UtahLakeZanazziS1$Collection_Date <- as.numeric(UtahLakeZanazziS1$Collection_Date)

#then turns collection data into data format, excel file did something weird where it was a weird number that represented number of days since dec. 30th 1899 so switched over to actual data format
UtahLakeZanazziS1$Collection_Date <- as.Date(UtahLakeZanazziS1$Collection_Date, origin = "1899-12-30")

```

```{r}

#adding date (20XX-XX)
UtahLakeZanazziS1 <- UtahLakeZanazziS1 %>%
  mutate(
    Date = format(Collection_Date, "%Y-%m"))

#adding Year column
UtahLakeZanazziS1 <- UtahLakeZanazziS1 %>%
  mutate(
    Year = format(Collection_Date, "%Y")
  )
#making year not a continuous value
UtahLakeZanazziS1$Year <- as.factor(UtahLakeZanazziS1$Year)

#making everything numeric, because everything started as a character value
UtahLakeZanazziS1$d18O <- as.numeric(UtahLakeZanazziS1$d18O)

UtahLakeZanazziS1$d2H <- as.numeric(UtahLakeZanazziS1$d2H)

UtahLakeZanazziS1$Latitude <- as.numeric(UtahLakeZanazziS1$Latitude)

UtahLakeZanazziS1$Longitude <- as.numeric(UtahLakeZanazziS1$Longitude)

UtahLakeZanazziS1$SD2H <- as.numeric(UtahLakeZanazziS1$SD2H)

UtahLakeZanazziS1$SD18O <- as.numeric(UtahLakeZanazziS1$SD18O)


#creating type row, this made it easier to readd to the utah county data
UtahLakeZanazziS1 <- UtahLakeZanazziS1 %>% 
  mutate(Type = "Zanazzi")

#creating DONOTUSE row
UtahLakeZanazziS1 <- UtahLakeZanazziS1 %>%
  mutate(
    DONOTUSE = FALSE
  )

```

#### Combining Zanazzi with water isotope

```{r}
#creating combined df, also added on row with row number so could get rid of unwanted values through donotuse and organization purposes
CombinedDF <- bind_rows(UtahLakeSamples, UtahLakeZanazziS1) %>% 
  mutate(
    Row = row_number()
  )
```

```{r}
#getting rid of far away sites or other weird #s, from map below 
CombinedDF <- CombinedDF %>% 
      mutate(DONOTUSE = Row %in% c(9, 19, 7, 18, 8, 17, 21 
          , TRUE, DONOTUSE))
```

##### Boxplots

```{r}

#Water Isotopes.org and zanazzi box plots to visualize the data in order to make decisions about analysis later
#d18O measurements 

#all data (includes ground water)
UtahLaked18Oplt <- UtahLakeSamples %>% 
  filter(DONOTUSE == FALSE) %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "All WI.org Samples"
  )+
  theme_classic()

UtahLaked18Oplt

#lake only samples
LakeOnlyd18Oplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond") %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "WI.org Lake Only Samples"
  )+
  #scale_y_continuous(limits = c(-17, -2))+
  theme_classic()

LakeOnlyd18Oplt


#Lake and river only
LakeandRiverd18Oplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream") %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "WI.org Lake and River Samples"
  )+
  #scale_y_continuous(limits = c(-17, -2))+
  theme_classic()

LakeandRiverd18Oplt


#Lake, River, Spring
LakeRiverSpringd18Oplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream" | Type == "Spring") %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "WI.org Lake, River, and Spring Samples"
  )+
  #scale_y_continuous(limits = c(-17, -2))+
  theme_classic()

LakeRiverSpringd18Oplt

#lake and spring only, pretty sure spring samples were just older lake samples
LakeSpringd18Oplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "Spring") %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "WI.org Lake and Spring Samples"
  )+
  #scale_y_continuous(limits = c(-17, -2))+
  theme_classic()

LakeSpringd18Oplt

#Zanazzi
#All samples
ZanazziSamplesd18Oplt <- UtahLakeZanazziS1 %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "All Zanazzi Samples"
  )+
  theme_minimal()

ZanazziSamplesd18Oplt

#all samples from zanazzi and water isotopes.org
CombinedSamplesd18Oplt <- CombinedDF %>% 
  filter(DONOTUSE == FALSE) %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "Collected Utah Lake Samples 1991 - 2024"
  )+
  theme_classic()

CombinedSamplesd18Oplt
```

```{r}
#saved last plot with all data including zanazzi for rachel's presentation
ggsave(
  "CombinedSamplesPlot3.pdf",
  plot = CombinedSamplesd18Oplt,
  width = 8,
  height = 6
)
```

```{r}
#Water Isotopes.org and zanazzi box plots to visualize the data in order to make decisions about analysis later
#d2H measurements

#lake only samples
LakeOnlyd2Hplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond") %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  theme_classic()

LakeOnlyd2Hplt


#Lake and river only
LakeandRiverd2Hplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream") %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  theme_classic()

LakeandRiverd2Hplt


#Lake, River, Spring
LakeRiverSpringd2Hplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream" | Type == "Spring") %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  theme_classic()

LakeRiverSpringd2Hplt

#lake and spring samples only
LakeSpringd2Hplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "Spring") %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  theme_classic()

LakeSpringd2Hplt

#all zanazzi samples
ZanazziSamplesd2Hplt <- UtahLakeZanazziS1 %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "All Zanazzi Samples"
  )+
  theme_minimal()

ZanazziSamplesd2Hplt

#combined samples
CombinedSamplesd2Hplt <- CombinedDF %>% 
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream" | Type == "Spring") %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "Collected Utah Lake Samples 1991 - 2024"
  )+
  theme_classic()

CombinedSamplesd2Hplt
```

##### Map Plot

```{r}
#used lat and long measurments to place them on a map and get rid of any samples that were not actually from utah lake, did this because originally all data from water isotopes.org was filtered from lat and long but therefore wasn't sure how close to utah lake the values actually were, this allowed me to see what was actually happening with values

#created dfs that were specific df and then converted them so they could be used on a map
#WI.org only
UtahLakeSamples_Filtered <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) 


UtahLakeSamples_sf <- st_as_sf(
  UtahLakeSamples_Filtered,
  coords = c("Longitude", "Latitude"),
             crs = 4326)
#zanazzi only
UtahLakeZanazzi_Filtered <- UtahLakeZanazziS1 %>%
  filter(DONOTUSE == FALSE) 


UtahLakeZanazzi_sf <- st_as_sf(
  UtahLakeZanazzi_Filtered,
  coords = c("Longitude", "Latitude"),
             crs = 4326)

#combined
UtahLakeCombined_Filtered <- CombinedDF %>%
  filter(DONOTUSE == FALSE) 


UtahLakeCombined_sf <- st_as_sf(
  UtahLakeCombined_Filtered,
  coords = c("Longitude", "Latitude"),
             crs = 4326)
```

```{r}
#had to create color pallettes for how i wanted to visualize the points placed on the map
#WI.org
#palette was based on year
pal1 <- colorFactor(
  palette = "Set2",
  domain = UtahLakeSamples_sf$Year
)

#zanazzi
#palette was based on which site it was
pal2 <- colorFactor(
  palette = "Set2",
  domain = UtahLakeZanazzi_sf$Site_Name
)

#combined
#palette was based on type of sample (either zanazzi, lake, river, or spring)
pal3 <- colorFactor(
  palette = "Set2",
  domain = UtahLakeCombined_sf$Type
)
```

```{r}
#actually creates the map
#WI.org
# Create shared data object
UtahLakeSharedSamples <- SharedData$new(UtahLakeSamples_sf)

# Interactive filtered map
bscols(
  widths = c(3, 9),
  
  filter_select(
    id = "Type",
    label = "Sample Type",
    sharedData = UtahLakeSharedSamples,
    group = ~Type
  ),
  
  leaflet(UtahLakeSharedSamples) %>%
    
    addProviderTiles(providers$Esri.WorldImagery) %>%
    
    addCircleMarkers(
      radius = 5,
      stroke = TRUE,
      weight = 1,
      fillOpacity = 0.8,
      
      color = ~pal1(Year),
      fillColor = ~pal1(Year),
      
      popup = ~paste(
        "Type:", Type,
        "<br>Year:", Year
      )
    ) %>%
    
    addLegend(
      position = "bottomright",
      pal = pal1,
      values = ~Year,
      title = "Year"
    ) %>%
    
    setView(
      lng = -111.85,
      lat = 40.25,
      zoom = 10
    )
)
```

```{r}
#Zanazzi
# Create shared data object
UtahLakeZanazziSharedSamples <- SharedData$new(UtahLakeZanazzi_sf)

# Interactive filtered map
bscols(
  widths = c(3, 9),
  
  filter_select(
    id = "Year",
    label = "Year",
    sharedData = UtahLakeZanazziSharedSamples,
    group = ~Year
  ),
  
  leaflet(UtahLakeZanazziSharedSamples) %>%
    
    addProviderTiles(providers$Esri.WorldImagery) %>%
    
    addCircleMarkers(
      radius = 5,
      stroke = TRUE,
      weight = 1,
      fillOpacity = 0.8,
      
      color = ~pal2(Site_Name),
      fillColor = ~pal2(Site_Name),
      
      popup = ~paste(
        "Site_Name:", Site_Name,
        "<br>Year:", Year
      )
    ) %>%
    
    addLegend(
      position = "bottomright",
      pal = pal2,
      values = ~Site_Name,
      title = "Site_Name"
    ) %>%
    
    setView(
      lng = -111.85,
      lat = 40.25,
      zoom = 10
    )
)
```

```{r}
#Combined
# Create shared data object
UtahLakeCombinedSharedSamples <- SharedData$new(UtahLakeCombined_sf)

# Interactive filtered map
bscols(
  widths = c(3, 9),
  
  filter_select(
    id = "Type",
    label = "Sample Type",
    sharedData = UtahLakeCombinedSharedSamples,
    group = ~Type
  ),
  
  leaflet(UtahLakeCombinedSharedSamples) %>%
    
    addProviderTiles(providers$Esri.WorldImagery) %>%
    
    addCircleMarkers(
      radius = 5,
      stroke = TRUE,
      weight = 1,
      fillOpacity = 0.8,
      
      color = ~pal3(Type),
      fillColor = ~pal3(Type),
      
      popup = ~paste(
        "Type:", Type,
        "<br>Year:", Year,
        "Row:", Row
      )
    ) %>%
    
    addLegend(
      position = "bottomright",
      pal = pal3,
      values = ~Type,
      title = "Type"
    ) %>%
    
    setView(
      lng = -111.85,
      lat = 40.25,
      zoom = 10
    )
)
```

##### d2H v. d18O

```{r}
#added d2H v. d18O plot to add to data comparison (didn't use in analysis decision as much)
#all samples
Alld2Hvd18Oplt <- UtahLakeSamples %>% 
  ggplot()+
  aes(x= d18O, y = d2H, shape = Type, color = Year)+
  geom_point()+
  theme_minimal()

Alld2Hvd18Oplt

#lake samples only
Laked2Hvd18Oplt <- UtahLakeSamples %>% 
  filter(Type == "Lake_or_pond") %>% 
  ggplot()+
  aes(x= d18O, y = d2H, shape = Type, color = Year)+
  geom_point()+
  theme_minimal()

Laked2Hvd18Oplt

#lake and river samples

LakeandRiverd2Hvd18Oplt <- UtahLakeSamples %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream") %>% 
  ggplot()+
  aes(x= d18O, y = d2H, shape = Type, color = Year)+
  geom_point()+
  theme_minimal()

LakeandRiverd2Hvd18Oplt

#Lake, River, and Spring

LakeRiverSpringd2Hvd18Oplt <- UtahLakeSamples %>%
  filter(Type == "Lake_or_pond" | Type == "River_or_stream" | Type == "Spring") %>% 
  ggplot()+
  aes(x= d18O, y = d2H, shape = Type, color = Year)+
  geom_point()+
  theme_minimal()

LakeRiverSpringd2Hvd18Oplt

#all samples
Combinedd2Hvd18Oplt <- CombinedDF %>% 
  filter(DONOTUSE == FALSE) %>% 
  ggplot()+
  aes(x = d18O, y = d2H, size = Year, color = Site_Name, shape = Type)+
  geom_point()+
  theme_classic()

ggplotly(Combinedd2Hvd18Oplt)
Combinedd2Hvd18Oplt
```

### Analysis

#### Precipitation (amount weighted)

$$\delta_w = \frac{\sum_{i=1}^{n} P_i \delta_i}{\sum_{i=1}^{n} P_i}$$

------------------------------------------------------------------------

##### SKIPPPPP

```{r}
#loads in amount of fallen precipitation in order to compare to precip isotopes and compute weighted amount
ONAQprecipitationamount <- loadByProduct(
  dpID = "DP1.00044.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE
)
```

```{r}
#creates df from precip amount for when measurments were taken, eithere every hour or once a day
ONAQPrecipAmountDaily <- as_tibble(ONAQprecipitationamount$WEIPRE_daily)

ONAQPrecipAmount60 <- as_tibble(ONAQprecipitationamount$WEIPRE_60min)
```

```{r}
#created csv from the two dfs above 
write_csv(ONAQPrecipAmount60, "ONAQPrecipAmount60.csv")

write_csv(ONAQPrecipAmountDaily, "ONAQPrecipAmountDaily.csv")
```

------------------------------------------------------------------------

##### START!

```{r}
#read the two csvs
ONAQPrecipAmount60 <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQPrecipAmount60.csv")

ONAQPrecipAmountDaily <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQPrecipAmountDaily.csv")
```

```{r}
#chose to mostly use daily measurments to deal with less data and it was more consistent with other data frames and other data averaging periods
#in daily precip measurments df - get rid of any days that have 0 precipitation
ONAQPrecipAmountCompare <- ONAQPrecipAmountDaily %>% 
  filter(precipBulk != 0)
```

```{r}
#renamed to be consistent
ONAQPrecipAmountCompare <- ONAQPrecipAmountCompare %>% 
  rename(
    "CollectionDate" = "date"
  )
```

```{r}
#Adding Date
ONAQPrecipAmountCompare <- ONAQPrecipAmountCompare %>% 
  mutate(
    Date = format(CollectionDate, "%Y-%m")
  )


```

```{r}
#calculating average precipitation amounts per month
ONAQPrecipAmountCompareAvg <- ONAQPrecipAmountCompare %>%
  group_by(Date) %>% 
  summarise(PrecipAmountAvg = mean(precipBulk, na.rm = TRUE))

#adding back in 2024-06
#originally did not exist so added as a 0 value
ONAQPrecipAmountCompareAvg %>%  add_row(Date = "2024-06", PrecipAmountAvg = 0)

```

##### Sample Plot:

```{r}
#wanted to visualize the precip amount over the time
ViewPrecipAmountplt <- ONAQPrecipAmountCompareAvg %>% 
  ggplot()+
  aes(x = Date, y = PrecipAmountAvg)+
  geom_point()+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45))

ViewPrecipAmountplt

ViewPrecipAmount2022plt <- ONAQPrecipAmountCompareAvg %>%
  filter(substr(Date, 1, 4) == "2022") %>% 
  ggplot()+
  aes(x = Date, y = PrecipAmountAvg)+
  geom_point()+
  theme_minimal()
  
ViewPrecipAmount2022plt


```

```{r}
#average each month of precipitation isotopes for d18O and d2H
#need to change to fit monte carlo analysis
ONAQPrecipMonthAvg <- ONAQPrecip %>%
  group_by(Date) %>% 
  summarise(PrecipAvgd18Omean = mean(d18O, na.rm = TRUE),
            PrecipAvgd2Hmean = mean(d2H, na.rm = TRUE))
```

```{r}
#adding in missing months to precip isotope values, same as above where the value was either zero or didn't exist so replaced with zero to keep meausrments consistent
ONAQPrecipMonthAvg %>% add_row(Date = "2022-02", PrecipAvgd18Omean = 0, PrecipAvgd2Hmean = 0)
ONAQPrecipMonthAvg %>% add_row(Date = "2024-06", PrecipAvgd18Omean = 0, PrecipAvgd2Hmean = 0)

#combining monthly averages of precip amounts and precip isotopes into the same df (d18O)
ONAQPrecipAmountd18OCombinedMonthAvg <-
  bind_cols(ONAQPrecipAmountCompareAvg %>% select(Date, PrecipAmountAvg), ONAQPrecipMonthAvg %>% select(PrecipAvgd18Omean)
  )

#combining monthly averages of precip amounts and precip isotopes into the same df (d2H)
ONAQPrecipAmountd2HCombinedMonthAvg <-
  bind_cols(ONAQPrecipAmountCompareAvg %>% select(Date, PrecipAmountAvg), ONAQPrecipMonthAvg %>% select(PrecipAvgd2Hmean)
  )
```

```{r}
#AGAIN CHANGE TO FIT MONTE CARLO
#calculating amount weighted isotopes for d18O
ONAQPrecipWeightedd18O <- ONAQPrecipAmountCombinedMonthAvg %>%
  group_by(Date) %>%
  summarise(
    d18OWeighted =
      sum(PrecipAmountAvg * PrecipAvgd18Omean, na.rm = TRUE) /
      sum(PrecipAmountAvg, na.rm = TRUE)
  )

#calculating amount weighted isotopes for d2H
ONAQPrecipWeightedd2H<- ONAQPrecipAmountCombinedMonthAvg %>%
  group_by(Date) %>%
  summarise(
    d2HWeighted =
      sum(PrecipAmountAvg * PrecipAvgd18Omean, na.rm = TRUE) /
      sum(PrecipAmountAvg, na.rm = TRUE)
  )
```

#### Fractionation Factor

$$\alpha = \frac{R_l}{R_v}$$

```{r}
#need to change so that it fits monte carlo analysis
#calculating and creating a new df for the monthly averages of atmosphere values for Rv value of fractionation factor equation 
ONAQAtmosMonthAvg <- ONAQAtmos %>%
  group_by(Month) %>%  
  summarise(AtmosAvgd18Omean = mean(d18Omean, na.rm = TRUE),
            AtmosAvgd2Hmean = mean(d2Hmean, na.rm = TRUE))

#same as previous chunk but creating montly averages for precipitation values, this would be Rl value of fraction factor equation
ONAQPrecipMonthAvg <- ONAQPrecip %>%
  group_by(Month) %>% 
  summarise(PrecipAvgd18Omean = mean(d18O, na.rm = TRUE),
            PrecipAvgd2Hmean = mean(d2H, na.rm = TRUE))

```

```{r}
#creating new df from precip montly average and atmosphere average with month and the averaged d18O value
ONAQd18OCombinedMonthAvg <- 
  bind_cols(ONAQPrecipMonthAvg %>% select(Month, PrecipAvgd18Omean), ONAQAtmosMonthAvg %>% select(AtmosAvgd18Omean)
  )

#creating new df from precip montly average and atmosphere average with month and the averaged d2H value
ONAQd2HCombinedMonthAvg <- 
  bind_cols(ONAQPrecipMonthAvg %>% select(Month, PrecipAvgd2Hmean), ONAQAtmosMonthAvg %>% select(AtmosAvgd2Hmean)
  )
```

```{r}
#above was incorrect because it didn't include the precip weighted values, created new df with the weighted values and the atmosphere values, only for d18O
ONAQd18OCombinedMonthAvg <- 
  bind_cols(Weightedd18O %>% select(Month, Weightedd18O), ONAQAtmosMonthAvg %>% select(AtmosAvgd18Omean)
  )

```

```{r}
#This is wrong!
#calcualted fractionation factor from precip and atmosphere, calcualted wrong fractionation factor
ONAQd18OCombinedMonthAvg <- ONAQd18OCombinedMonthAvg %>% 
  mutate(
    FractionationFactor = (Weightedd18O/AtmosAvgd18Omean))

ONAQd2HCombinedMonthAvg <- ONAQd2HCombinedMonthAvg %>% 
  mutate(
    FractionationFactor = (PrecipAvgd2Hmean/AtmosAvgd2Hmean))
  
```

### Plots

a few random plots that were created at the very beginning to visualize data and see what i was working with and what was happening in the data

```{r}
ONAQAtmplt1 <- ONAQAtmos %>% 
  ggplot()+
  aes(x = Year, y = d18Omean)+
  geom_boxplot(fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  theme_classic()

ONAQAtmplt1
```

```{r}
ONAQAtmplt2 <- ONAQAtmos %>% 
  ggplot()+
  aes(x = d18Omean, y = d2Hmean)+
  geom_point()+
  geom_abline(slope = 7, intercept = 0, color = "#6FA458", linewidth = 1) +
  geom_abline(slope = 7.45, intercept = -1.65, color = "#5B2C4D", linetype = 3, size = 2)+
  theme_classic()

ONAQAtmplt2
```

```{r}
ONAQAtm2022 <- ONAQAtmos %>% 
  filter(Year == 2022)  


ONAQAtmHumidity2022 <- ONAQAtm2022 %>% 
  ggplot()+
  aes(x = StartTime,  y = humidityratio)+
  geom_point()+
  theme_classic()

ONAQAtmHumidity2022
```

```{r}
ONAQd18OCombinedMonthAvgPlt <- ONAQd18OCombinedMonthAvg %>% 
  ggplot()+
  geom_point(aes(x = Month, y = AtmosAvgd18Omean))+
  geom_point(aes(x = Month, y = PrecipAvgd18Omean))+
  theme_classic()

ONAQd18OCombinedMonthAvgPlt
```
