---
title: "UROPSum_UnneededCode"
format: html
editor: visual
---

## April 2024 Test Run

```{r}
#april 2024 atmosphere data, test run so could deal with smaller data set. this generates the #h5 files on the computer but need to be picked apart manually on the computer manually
zipsByProduct(
  dpID = "DP4.00200.001",
  site = "ONAQ",
  startdate = "2024-04",
  enddate = "2024-04",
  package = "basic",
  check.size = FALSE
)
```

```{r}
#stacks the files to be a df in R from the manually unstacked h5 files. file path leads to the h5 files that i manually placed into this folder
AprilEddy <- stackEddy(
  filepath = "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/filesToStack00200/AprilSampleh5",
  level = "dp01",
  var = "tempAirTop",
  avg = 30
)

```

```{r}
#creates april 2024 only atmosphere isotopes csv file
write.csv(Eddy$ONAQ, file = "ONAQAprilAtm.csv")

```

```{r}
#reads april 2024 only atmosphere isotopes csv file
ONAQTRUEAprilAtm <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/ONAQAprilAtm.csv")
```

## Humidity Mole Ratio

```{r}
#originally intended to calculate humidity here but instead chose to use the actual data set from NEON tracking humidity so decided to not use this caluclation and voided it from the code
#ONAQAtmos <- ONAQAtmos %>% 
  #mutate(
    #humidityratio = ((WetMoleRatioMean / DryMoleRatioMean) * 0.677)
  #)
```

## Humidity checks originally sperated by different towers, but changed to averaging all sensors

## 

```{r}
#data frame for 40 ft tower measurment of humidity
ONAQHumid40 <- ONAQHumid %>% 
  filter(verticalPosition == "040")

#data frame for 0ft tower measurment of humidity
ONAQHumid0 <- ONAQHumid %>% 
  filter(verticalPosition == "000")
```

# Checks

```{r}
#created a data frame checking the vertical position variable in the atmosphere df and what the different positions were and their effect on if the data was NA or not
#vertcheck <- ONAQAtmos %>% 
  #filter(verticalPosition == "h2oMed")

#vertcheck <- vertcheck %>% 
  #filter(!is.na(d18Omean))

```

## weird zanazzi files

```{r}
#reads each of these files of individual sheets, but only need sheet 1 because these are all repeats individual collections
UtahLakeZanazziS2 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/Table S2.csv")

UtahLakeZanazziS3 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/Table S3.csv")

UtahLakeZanazziS4 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/Table S4.csv")

UtahLakeZanazziS5 <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/Table S5.csv")
```

## Utah Lake Isotope Data Plots 

### Boxplots

```{r}

#Water Isotopes.org and zanazzi box plots to visualize the data in order to make decisions about analysis later
#d18O measurements 

#all data (includes ground water)
UtahLaked18Oplt <- UtahLakeSamples %>% 
  filter(DONOTUSE == FALSE) %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "All WI.org Samples"
  )+
  theme_classic()

UtahLaked18Oplt

#lake only samples
LakeOnlyd18Oplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond") %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "WI.org Lake Only Samples"
  )+
  #scale_y_continuous(limits = c(-17, -2))+
  theme_classic()

LakeOnlyd18Oplt


#Lake and river only
LakeandRiverd18Oplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream") %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "WI.org Lake and River Samples"
  )+
  #scale_y_continuous(limits = c(-17, -2))+
  theme_classic()

LakeandRiverd18Oplt


#Lake, River, Spring
LakeRiverSpringd18Oplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream" | Type == "Spring") %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "WI.org Lake, River, and Spring Samples"
  )+
  #scale_y_continuous(limits = c(-17, -2))+
  theme_classic()

LakeRiverSpringd18Oplt

#lake and spring only, pretty sure spring samples were just older lake samples
LakeSpringd18Oplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "Spring") %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "WI.org Lake and Spring Samples"
  )+
  #scale_y_continuous(limits = c(-17, -2))+
  theme_classic()

LakeSpringd18Oplt

#Zanazzi
#All samples
ZanazziSamplesd18Oplt <- UtahLakeZanazziS1 %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "All Zanazzi Samples"
  )+
  theme_minimal()

ZanazziSamplesd18Oplt

#all samples from zanazzi and water isotopes.org
CombinedSamplesd18Oplt <- CombinedDF %>% 
  filter(DONOTUSE == FALSE) %>% 
  ggplot()+
  aes(x = Year, y = d18O)+
  geom_boxplot(
    fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "Collected Utah Lake Samples 1991 - 2024"
  )+
  theme_classic()

CombinedSamplesd18Oplt
```

```{r}
#saved last plot with all data including zanazzi for rachel's presentation
ggsave(
  "CombinedSamplesPlot3.pdf",
  plot = CombinedSamplesd18Oplt,
  width = 8,
  height = 6
)
```

```{r}
#Water Isotopes.org and zanazzi box plots to visualize the data in order to make decisions about analysis later
#d2H measurements

#lake only samples
LakeOnlyd2Hplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond") %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  theme_classic()

LakeOnlyd2Hplt


#Lake and river only
LakeandRiverd2Hplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream") %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  theme_classic()

LakeandRiverd2Hplt


#Lake, River, Spring
LakeRiverSpringd2Hplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream" | Type == "Spring") %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  theme_classic()

LakeRiverSpringd2Hplt

#lake and spring samples only
LakeSpringd2Hplt <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "Spring") %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  theme_classic()

LakeSpringd2Hplt

#all zanazzi samples
ZanazziSamplesd2Hplt <- UtahLakeZanazziS1 %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "All Zanazzi Samples"
  )+
  theme_minimal()

ZanazziSamplesd2Hplt

#combined samples
CombinedSamplesd2Hplt <- CombinedDF %>% 
  filter(DONOTUSE == FALSE) %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream" | Type == "Spring") %>% 
  ggplot()+
  aes(x = Year, y = d2H)+
  geom_boxplot(
    fill = "indianred", median.linewidth = 1.0, median.color = "indianred4", box.color = "black", whisker.color = "black"
  )+
  labs(
    title = "Collected Utah Lake Samples 1991 - 2024"
  )+
  theme_classic()

CombinedSamplesd2Hplt
```

### Map Plot

```{r}
#used lat and long measurments to place them on a map and get rid of any samples that were not actually from utah lake, did this because originally all data from water isotopes.org was filtered from lat and long but therefore wasn't sure how close to utah lake the values actually were, this allowed me to see what was actually happening with values

#created dfs that were specific df and then converted them so they could be used on a map
#WI.org only
UtahLakeSamples_Filtered <- UtahLakeSamples %>%
  filter(DONOTUSE == FALSE) 


UtahLakeSamples_sf <- st_as_sf(
  UtahLakeSamples_Filtered,
  coords = c("Longitude", "Latitude"),
             crs = 4326)
#zanazzi only
UtahLakeZanazzi_Filtered <- UtahLakeZanazziS1 %>%
  filter(DONOTUSE == FALSE) 


UtahLakeZanazzi_sf <- st_as_sf(
  UtahLakeZanazzi_Filtered,
  coords = c("Longitude", "Latitude"),
             crs = 4326)

#combined
UtahLakeCombined_Filtered <- CombinedDF %>%
  filter(DONOTUSE == FALSE) 


UtahLakeCombined_sf <- st_as_sf(
  UtahLakeCombined_Filtered,
  coords = c("Longitude", "Latitude"),
             crs = 4326)
```

```{r}
#had to create color pallettes for how i wanted to visualize the points placed on the map
#WI.org
#palette was based on year
pal1 <- colorFactor(
  palette = "Set2",
  domain = UtahLakeSamples_sf$Year
)

#zanazzi
#palette was based on which site it was
pal2 <- colorFactor(
  palette = "Set2",
  domain = UtahLakeZanazzi_sf$Site_Name
)

#combined
#palette was based on type of sample (either zanazzi, lake, river, or spring)
pal3 <- colorFactor(
  palette = "Set2",
  domain = UtahLakeCombined_sf$Type
)
```

```{r}
#actually creates the map
#WI.org
# Create shared data object
UtahLakeSharedSamples <- SharedData$new(UtahLakeSamples_sf)

# Interactive filtered map
bscols(
  widths = c(3, 9),
  
  filter_select(
    id = "Type",
    label = "Sample Type",
    sharedData = UtahLakeSharedSamples,
    group = ~Type
  ),
  
  leaflet(UtahLakeSharedSamples) %>%
    
    addProviderTiles(providers$Esri.WorldImagery) %>%
    
    addCircleMarkers(
      radius = 5,
      stroke = TRUE,
      weight = 1,
      fillOpacity = 0.8,
      
      color = ~pal1(Year),
      fillColor = ~pal1(Year),
      
      popup = ~paste(
        "Type:", Type,
        "<br>Year:", Year
      )
    ) %>%
    
    addLegend(
      position = "bottomright",
      pal = pal1,
      values = ~Year,
      title = "Year"
    ) %>%
    
    setView(
      lng = -111.85,
      lat = 40.25,
      zoom = 10
    )
)
```

```{r}
#Zanazzi
# Create shared data object
UtahLakeZanazziSharedSamples <- SharedData$new(UtahLakeZanazzi_sf)

# Interactive filtered map
bscols(
  widths = c(3, 9),
  
  filter_select(
    id = "Year",
    label = "Year",
    sharedData = UtahLakeZanazziSharedSamples,
    group = ~Year
  ),
  
  leaflet(UtahLakeZanazziSharedSamples) %>%
    
    addProviderTiles(providers$Esri.WorldImagery) %>%
    
    addCircleMarkers(
      radius = 5,
      stroke = TRUE,
      weight = 1,
      fillOpacity = 0.8,
      
      color = ~pal2(Site_Name),
      fillColor = ~pal2(Site_Name),
      
      popup = ~paste(
        "Site_Name:", Site_Name,
        "<br>Year:", Year
      )
    ) %>%
    
    addLegend(
      position = "bottomright",
      pal = pal2,
      values = ~Site_Name,
      title = "Site_Name"
    ) %>%
    
    setView(
      lng = -111.85,
      lat = 40.25,
      zoom = 10
    )
)
```

```{r}
#Combined
# Create shared data object
UtahLakeCombinedSharedSamples <- SharedData$new(UtahLakeCombined_sf)

# Interactive filtered map
bscols(
  widths = c(3, 9),
  
  filter_select(
    id = "Type",
    label = "Sample Type",
    sharedData = UtahLakeCombinedSharedSamples,
    group = ~Type
  ),
  
  leaflet(UtahLakeCombinedSharedSamples) %>%
    
    addProviderTiles(providers$Esri.WorldImagery) %>%
    
    addCircleMarkers(
      radius = 5,
      stroke = TRUE,
      weight = 1,
      fillOpacity = 0.8,
      
      color = ~pal3(Type),
      fillColor = ~pal3(Type),
      
      popup = ~paste(
        "Type:", Type,
        "<br>Year:", Year,
        "Row:", Row
      )
    ) %>%
    
    addLegend(
      position = "bottomright",
      pal = pal3,
      values = ~Type,
      title = "Type"
    ) %>%
    
    setView(
      lng = -111.85,
      lat = 40.25,
      zoom = 10
    )
)
```

### d2H v. d18O

```{r}
#added d2H v. d18O plot to add to data comparison (didn't use in analysis decision as much)
#all samples
Alld2Hvd18Oplt <- UtahLakeSamples %>% 
  ggplot()+
  aes(x= d18O, y = d2H, shape = Type, color = Year)+
  geom_point()+
  theme_minimal()

Alld2Hvd18Oplt

#lake samples only
Laked2Hvd18Oplt <- UtahLakeSamples %>% 
  filter(Type == "Lake_or_pond") %>% 
  ggplot()+
  aes(x= d18O, y = d2H, shape = Type, color = Year)+
  geom_point()+
  theme_minimal()

Laked2Hvd18Oplt

#lake and river samples

LakeandRiverd2Hvd18Oplt <- UtahLakeSamples %>% 
  filter(Type == "Lake_or_pond" | Type == "River_or_stream") %>% 
  ggplot()+
  aes(x= d18O, y = d2H, shape = Type, color = Year)+
  geom_point()+
  theme_minimal()

LakeandRiverd2Hvd18Oplt

#Lake, River, and Spring

LakeRiverSpringd2Hvd18Oplt <- UtahLakeSamples %>%
  filter(Type == "Lake_or_pond" | Type == "River_or_stream" | Type == "Spring") %>% 
  ggplot()+
  aes(x= d18O, y = d2H, shape = Type, color = Year)+
  geom_point()+
  theme_minimal()

LakeRiverSpringd2Hvd18Oplt

#all samples
Combinedd2Hvd18Oplt <- CombinedDF %>% 
  filter(DONOTUSE == FALSE) %>% 
  ggplot()+
  aes(x = d18O, y = d2H, size = Year, color = Site_Name, shape = Type)+
  geom_point()+
  theme_classic()

ggplotly(Combinedd2Hvd18Oplt)
Combinedd2Hvd18Oplt
```

## Visualizing precip amounts

```{r}
#wanted to visualize the precip amount over the time
ViewPrecipAmountplt <- ONAQPrecipAmountCompareAvg %>% 
  ggplot()+
  aes(x = Date, y = PrecipAmountAvg)+
  geom_point()+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45))

ViewPrecipAmountplt

ViewPrecipAmount2022plt <- ONAQPrecipAmountCompareAvg %>%
  filter(substr(Date, 1, 4) == "2022") %>% 
  ggplot()+
  aes(x = Date, y = PrecipAmountAvg)+
  geom_point()+
  theme_minimal()
  
ViewPrecipAmount2022plt


```

## Daytime v. Nighttime Code

```{r}
SoilAvgTemp <- SoilAvgTemp %>% 
  left_join(
    Daylight %>% select(startDateTime, Daylight),
    by = "startDateTime"
  )

# Sunrise_Sunset <- SoilRadiation %>%
#   mutate(
#     daylight = dirRadMean > 10
#   ) %>%
#   filter(daylight) %>%
#   group_by(Date) %>%
#   summarise(
#     Sunrise = first(Time),
#     Sunset = last(Time),
#     .groups = "drop"
#   )

Daylight <- SoilRadiation %>% 
  select(Daylight, endDateTime, startDateTime)

```

# Analysis that isn't monte carlo

```{r}
#average each month of precipitation isotopes for d18O and d2H
#need to change to fit monte carlo analysis
ONAQPrecipMonthAvg <- ONAQPrecip %>%
  group_by(Date) %>% 
  summarise(PrecipAvgd18Omean = mean(d18O, na.rm = TRUE),
            PrecipAvgd2Hmean = mean(d2H, na.rm = TRUE))
```

```{r}
#adding in missing months to precip isotope values, same as above where the value was either zero or didn't exist so replaced with zero to keep meausrments consistent
ONAQPrecipMonthAvg %>% add_row(Date = "2022-02", PrecipAvgd18Omean = 0, PrecipAvgd2Hmean = 0)
ONAQPrecipMonthAvg %>% add_row(Date = "2024-06", PrecipAvgd18Omean = 0, PrecipAvgd2Hmean = 0)

#combining monthly averages of precip amounts and precip isotopes into the same df (d18O)
ONAQPrecipAmountd18OCombinedMonthAvg <-
  bind_cols(ONAQPrecipAmountCompareAvg %>% select(Date, PrecipAmountAvg), ONAQPrecipMonthAvg %>% select(PrecipAvgd18Omean)
  )

#combining monthly averages of precip amounts and precip isotopes into the same df (d2H)
ONAQPrecipAmountd2HCombinedMonthAvg <-
  bind_cols(ONAQPrecipAmountCompareAvg %>% select(Date, PrecipAmountAvg), ONAQPrecipMonthAvg %>% select(PrecipAvgd2Hmean)
  )
```

```{r}
#AGAIN CHANGE TO FIT MONTE CARLO
#calculating amount weighted isotopes for d18O
ONAQPrecipWeightedd18O <- ONAQPrecipAmountCombinedMonthAvg %>%
  group_by(Date) %>%
  summarise(
    d18OWeighted =
      sum(PrecipAmountAvg * PrecipAvgd18Omean, na.rm = TRUE) /
      sum(PrecipAmountAvg, na.rm = TRUE)
  )

#calculating amount weighted isotopes for d2H
ONAQPrecipWeightedd2H<- ONAQPrecipAmountCombinedMonthAvg %>%
  group_by(Date) %>%
  summarise(
    d2HWeighted =
      sum(PrecipAmountAvg * PrecipAvgd18Omean, na.rm = TRUE) /
      sum(PrecipAmountAvg, na.rm = TRUE)
  )
```

## Fractionation Factor

$$\alpha = \frac{R_l}{R_v}$$

```{r}
#need to change so that it fits monte carlo analysis
#calculating and creating a new df for the monthly averages of atmosphere values for Rv value of fractionation factor equation 
ONAQAtmosMonthAvg <- ONAQAtmos %>%
  group_by(Month) %>%  
  summarise(AtmosAvgd18Omean = mean(d18Omean, na.rm = TRUE),
            AtmosAvgd2Hmean = mean(d2Hmean, na.rm = TRUE))

#same as previous chunk but creating montly averages for precipitation values, this would be Rl value of fraction factor equation
ONAQPrecipMonthAvg <- ONAQPrecip %>%
  group_by(Month) %>% 
  summarise(PrecipAvgd18Omean = mean(d18O, na.rm = TRUE),
            PrecipAvgd2Hmean = mean(d2H, na.rm = TRUE))

```

```{r}
#creating new df from precip montly average and atmosphere average with month and the averaged d18O value
ONAQd18OCombinedMonthAvg <- 
  bind_cols(ONAQPrecipMonthAvg %>% select(Month, PrecipAvgd18Omean), ONAQAtmosMonthAvg %>% select(AtmosAvgd18Omean)
  )

#creating new df from precip montly average and atmosphere average with month and the averaged d2H value
ONAQd2HCombinedMonthAvg <- 
  bind_cols(ONAQPrecipMonthAvg %>% select(Month, PrecipAvgd2Hmean), ONAQAtmosMonthAvg %>% select(AtmosAvgd2Hmean)
  )
```

```{r}
#above was incorrect because it didn't include the precip weighted values, created new df with the weighted values and the atmosphere values, only for d18O
ONAQd18OCombinedMonthAvg <- 
  bind_cols(Weightedd18O %>% select(Month, Weightedd18O), ONAQAtmosMonthAvg %>% select(AtmosAvgd18Omean)
  )

```

```{r}
#This is wrong!
#calcualted fractionation factor from precip and atmosphere, calcualted wrong fractionation factor
ONAQd18OCombinedMonthAvg <- ONAQd18OCombinedMonthAvg %>% 
  mutate(
    FractionationFactor = (Weightedd18O/AtmosAvgd18Omean))

ONAQd2HCombinedMonthAvg <- ONAQd2HCombinedMonthAvg %>% 
  mutate(
    FractionationFactor = (PrecipAvgd2Hmean/AtmosAvgd2Hmean))
  
```

# visualization plots

### Plots

a few random plots that were created at the very beginning to visualize data and see what i was working with and what was happening in the data

```{r}
ONAQAtmplt1 <- ONAQAtmos %>% 
  ggplot()+
  aes(x = Year, y = d18Omean)+
  geom_boxplot(fill = "lightblue", median.linewidth = 1.0, median.color = "lightblue4", box.color = "black", whisker.color = "black"
  )+
  theme_classic()

ONAQAtmplt1
```

```{r}
ONAQAtmplt2 <- ONAQAtmos %>% 
  ggplot()+
  aes(x = d18Omean, y = d2Hmean)+
  geom_point()+
  geom_abline(slope = 7, intercept = 0, color = "#6FA458", linewidth = 1) +
  geom_abline(slope = 7.45, intercept = -1.65, color = "#5B2C4D", linetype = 3, size = 2)+
  theme_classic()

ONAQAtmplt2
```

```{r}
ONAQAtm2022 <- ONAQAtmos %>% 
  filter(Year == 2022)  


ONAQAtmHumidity2022 <- ONAQAtm2022 %>% 
  ggplot()+
  aes(x = StartTime,  y = humidityratio)+
  geom_point()+
  theme_classic()

ONAQAtmHumidity2022
```

```{r}
ONAQd18OCombinedMonthAvgPlt <- ONAQd18OCombinedMonthAvg %>% 
  ggplot()+
  geom_point(aes(x = Month, y = AtmosAvgd18Omean))+
  geom_point(aes(x = Month, y = PrecipAvgd18Omean))+
  theme_classic()

ONAQd18OCombinedMonthAvgPlt
```

### dS and dT visualization plots

```{r}
dS2Hplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d2HdS)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dS2Hplt

dT2Hplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d2HdT)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dT2Hplt

dSd18Oplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d18OdS)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dSd18Oplt

dTd18Oplt <- ONAQdSdT %>% 
  ggplot()+
  aes(x = SampleDate, y = d18OdT)+
  geom_point()+
  geom_smooth()+
  theme_minimal()

dTd18Oplt
```
