---
title: "UROPSummer2026"
format: html
editor: visual
---

```{r}
#did this so it would use files from UROP Summer 2026 but doesn't work and files to still automatically go to the git hub file, therefore need to specify when calling the file in individual code lines

#don't need to run for code to work
setwd("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026")
```

```{r}
library(tidyverse)
library(latex2exp)
library(readxl)
library(egg)
library(plotly)
library(readr)
library(neonUtilities) #download and manage NEON data (in-house)
library(neonOS) #NEON OS data
library(terra) #general geological data
library(ggplot2)
library(colorspace)#got for using more colors, don't really use
library(viridis)#virids color palette
library(sf)#need for interactive map
library(tigris)#interactive map
library(leaflet)#need for satellite image of interactive map
library(maps)#interactive map
library(crosstalk)
```

# Downloading and Creating

## Atm and Air Temp

```{r}
#same as april 2024 above. pulls out zip file of all eddy covalence data. need to go in #manually and spereate out everything into its own file with all the h5 files.
zipsByProduct(
  dpID = "DP4.00200.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE
)
```

```{r}
#pulls out atmosphere isotopes from eddy covalence data and puts it into its own df inside R
AtmosEddy <- stackEddy(
  filepath = "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQAtmosH5",
  level = "dp01",
  var = "isoH2o",
  avg = 30
)

#does the same thing as the atmosphere, pulls out air temp from eddy covalence data and then organizes into its own df
AirTempEddy <- stackEddy(
  filepath = "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQAtmosH5",
  level = "dp01",
  var = "tempAirTop",
  avg = 30
)
```

```{r}
#writes the csv file for both atmosphere isotopes data and air temp data from eddy covalence from the df made above from the stack eddy func. these now exsit on the computer
write_csv(AtmosEddy$ONAQ, file = "ONAQAtmos.csv")
write_csv(AirTempEddy$ONAQ, file = "ONAQAirTemp.csv")
```

## Soil Water Iso

```{r}
#pulls data from the soil data from NEON. did the same thing where it stacked together all of the zip files and had to back and manually pull them out. but this isn't the eddy covalence, its for the soil data on neon
zipsByProduct(
  dpID = "DP1.10086.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE
  )
```

```{r}
#pulls soil isotope data from the h5 files and creates new df
SoilIsoEddy <- stackEddy(
  filepath = "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/filesToStack10086",
  level = "dp01",
  var = "soil",
  avg = 30
)
```

Need to add in csv file for this \*maybe

couldn't find the actual soil isotope data for this. probably don't need for utah lake calculations. might need for matos but they use their own df which is found in their example code file

## Humidity

```{r}
#load humidity df
ONAQhumidity <- loadByProduct(
  dpID = "DP1.00098.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE
  
)

#as tibble function creates own df with the 30 min measurments then the write csv creates the specific csv file in urop sum 2026 file
ONAQHumid <- as_tibble(ONAQhumidity$RH_30min)
write_csv(ONAQHumid, "ONAQRHumid.csv")
```

## Precipitation

```{r}
#same as humidity above, loads data from cloud/website into the r environment
ONAQprecipitation <- loadByProduct(dpID = "DP1.00038.001", 
                                   site = "ONAQ",
                                   startdate = "2022-01",
                                   enddate = "2024-12",
                                   check.size = FALSE)

#now creates df with just the precipitation isotope measurements, and then turns this into a csv file that will now exist on the computer
ONAQPrecip <- as_tibble(ONAQprecipitation$wdi_isoPerSample)
write_csv(ONAQPrecip, "ONAQPrecip.csv")
```

## Amount of Precip

```{r}
#loads in amount of fallen precipitation in order to compare to precip isotopes and compute weighted amount
ONAQprecipitationamount <- loadByProduct(
  dpID = "DP1.00044.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE
)
```

```{r}
#creates df from precip amount for when measurments were taken, eithere every hour or once a day
ONAQPrecipAmountDaily <- as_tibble(ONAQprecipitationamount$WEIPRE_daily)
ONAQPrecipAmount60 <- as_tibble(ONAQprecipitationamount$WEIPRE_60min)

#created csv from the two dfs above 
write_csv(ONAQPrecipAmount60, "ONAQPrecipAmount60.csv")
write_csv(ONAQPrecipAmountDaily, "ONAQPrecipAmountDaily.csv")
```

## Soil Radiation

```{r}
# soil radiation
SoilRadiation <- loadByProduct(
  dpID = "DP1.00014.001",
  site = "ONAQ",
  startdate = "2022-01",
  enddate = "2024-12",
  package = "basic",
  check.size = FALSE)

write_csv(SoilRadiation$SRDDP_30min, "SoilRadiation.csv")
```

```{r}
#loading in the data from zanazzi et al. 2020 
UtahLakeZanazziPath <- "/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/UtahLake2014-2016Zanazzi.xlsx"

#this df was weird and had lots of sheets inside the actaul file, so had to seperate out all the sheets in order to use the one i wanted
#names object sheet that lists all of the sheets in the file
sheets <- excel_sheets(UtahLakeZanazziPath)

#creates individual files for each sheet in the file
for(sheet in sheets) {
  
  UtahLakeZanazziSheet <- read_excel(UtahLakeZanazziPath, sheet = sheet)
  csv_name <- paste0(sheet, ".csv")
  write.csv(UtahLakeZanazziSheet, csv_name, row.names = FALSE)
}

```

# Start

## Load csv files in

```{r}
#loads in csv file for onaq atmospheric isotopes (again have to include the whole string because of weird file path stuff), TRUE label is not edited at all and is straight original version
ONAQAtmos <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQAtmos.csv")

ONAQPrecip <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQPrecip.csv")

ONAQHumid <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQRHumid.csv")

ONAQPrecipAmount <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/ONAQPrecipAmountDaily.csv")

SoilRadiation <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SoilRadiation.csv")

SingleAirTemp <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/SingleAirTemp.csv")

TripleAirTemp <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/TripleAirTemp.csv")

#data from water isotopes.org, no date restrictions, no type restrictions, restricted by lat and long to only get utah county data - date range is 1990 - 2024
#wanted this data to get a picture of all data collected and then sample from this later for monte carlo
UtahLakeCounty <- read_csv("/Users/haydenbosshardt/Desktop/SRI/RFiles/UROP/UROPSummer2026/AllUtahCountyData.csv")

#Zanazzi et al 2020 data, combination of all collected data so only read this csv file
UtahLakeZanazzi <- read_csv("~/Desktop/SRI/RFiles/UROP/UROPSummer2026/Table S1.csv")
```

```{r}
#creating function for averaging periods based on how many days you want to average from, need to put in which data frame to chose from, which value you are averaging, and the number of days in the averaging period
#this is what matos et al. decided to do, however, not sure if want this for the Utah lake data, because they were dealing with a lot more data than this project and it might create a more accurate representation to average for either one day or not at all

#calculates the mean, standard deviation, how many rows it averaged and start and end date
create_period_summary <- function(data,
                                  value_column,
                                  period_days) {
  start_date <- min(data$Date, na.rm = TRUE)
  data %>%
    mutate(
      PeriodGroup =
        floor(as.numeric(Date - start_date) / period_days)
    ) %>%
    group_by(PeriodGroup) %>%
    summarize(
      startDateTime = min(Date),
      endDateTime = max(Date),
      Mean = mean(.data[[value_column]], na.rm = TRUE),
      SD = sd(.data[[value_column]], na.rm = TRUE),
      N = sum(!is.na(.data[[value_column]])),
      .groups = "drop"
    )
}
```

# Data Frame Processing

## Atmosphere

```{r}
#modifies names of columns for easier work later and ease of understanding the data
ONAQAtmos <- ONAQAtmos %>% 
  rename(
    "d18Omean" = "data.isoH2o.dlta18OH2o.mean",
    "d2Hmean" = "data.isoH2o.dlta2HH2o.mean",
    "startDateTime" = "timeBgn",
    "endDateTime" = "timeEnd"
  )

#getting rid of unnecssary columns, specifically issue logs or repeats of data columns that came in a different format
ONAQAtmos <- ONAQAtmos %>% 
  select(horizontalPosition, verticalPosition, startDateTime, endDateTime, d18Omean, d2Hmean)


#getting rid of NA values

#gets rid of any NA with d18O isotope measurments as extra precaution
ONAQAtmos <- ONAQAtmos %>% 
  filter(!is.na(d18Omean))

#gets rid of any NA with d2H isotope measurments as extra precaution
ONAQAtmos <- ONAQAtmos %>% 
  filter(!is.na(d2Hmean))

```

```{r}
#averaging by startDateTime so you get averages of horizontal and vertical position measurments
ONAQAtmos <- ONAQAtmos %>% 
  group_by(startDateTime) %>%
  summarise(
    d18Omean = mean(d18Omean, na.rm = TRUE),
    d2Hmean = mean(d2Hmean, na.rm = TRUE),
    .groups = "drop"
  )

#added in more accurate date column
ONAQAtmos <- ONAQAtmos %>% 
  mutate(
    Date = as.Date(startDateTime)) 

#adding row numbers
ONAQAtmos <- ONAQAtmos %>% 
  mutate(
    Row = row_number()
  )

```

## Precipitation

```{r}
#renamed columns to be easier to use and understand
ONAQPrecip <- ONAQPrecip %>% 
  rename(
    "d18O" = "d18OWater",
    "d2H" = "d2HWater",
    "startDateTime" = "collectDate"
  )

#getting rid of unwanted columns (stuff to do with issue log or other variables not needed)
ONAQPrecip <-  ONAQPrecip %>% 
  select(startDateTime, d18O, d2H, startDateTime, d18OWaterUncert, d2HWaterUncert)

#added in only date column
ONAQPrecip <- ONAQPrecip %>% 
  mutate(
    Date = as.Date(startDateTime))

#adding row numbers
ONAQPrecip <- ONAQPrecip %>% 
  mutate(
    Row = row_number()
  )

#didn't add second averaging function because only one sensor
```

## Humidity

```{r}
#rename column names for easier use and understanding, most didn't need to be renamed because already easy to use and in a format that was consistent
ONAQHumid <- ONAQHumid %>% 
  rename(
    "MeanTempC" = "tempRHMean",
  )


#average humidity measurments from multiple different sensors by start of collection time, sensed from different heights so averaged to be one measurment
ONAQHumid <- ONAQHumid %>% 
  group_by(startDateTime) %>%
  summarise(
    RHMean = mean(RHMean, na.rm = TRUE),
    dewTempMean = mean(dewTempMean, na.rm = TRUE),
    .groups = "drop"
  )

#added in more accurate date time column
ONAQHumid <- ONAQHumid %>% 
  mutate(
    Date = as.Date(startDateTime)) 

#adding row numbers
ONAQHumid <- ONAQHumid %>% 
  mutate(
    Row = row_number()
  )
```

## Amount Precip 

```{r}
#renamed to be consistent
ONAQPrecipAmount <- ONAQPrecipAmount %>% 
  rename(
    "startDateTime" = "date"
  )

ONAQPrecipAmount <- ONAQPrecipAmount %>% 
  group_by(startDateTime) %>%
  summarise(
    precipBulk = mean(precipBulk, na.rm = TRUE),
    precipBulkExpUncert = mean(precipBulkExpUncert, na.rm = TRUE),
    .groups = "drop"
  )

#added in more accurate date time column
ONAQPrecipAmount <- ONAQPrecipAmount %>% 
  mutate(
    Date = as.Date(startDateTime))

#adding in row numbers
ONAQPrecipAmount <- ONAQPrecipAmount %>% 
  mutate(
    Row = row_number()
  )
```

## Soil Radiation

##### Identifying Sunset v. Sunrise

might not need if dealing with lake evaporation which does not fluctuate too muc and fluctuations can be taken into consideration with either an average or with monte carlo analysis

```{r}

SoilRadiation <- SoilRadiation %>% 
  group_by(startDateTime) %>%
  summarise(
    difRadMean = mean(difRadMean, na.rm = TRUE),
    dirRadMean = mean(dirRadMean, na.rm = TRUE),
    gloRadMean = mean(gloRadMean, na.rm = TRUE),
    sunPres = mean(sunPres, na.rm = TRUE),
    .groups = "drop"
  )

#adding in daylight column - might not need/use
SoilRadiation <- SoilRadiation %>% 
  mutate(
    Daylight = difRadMean > 10
  )

#making date column
SoilRadiation <- SoilRadiation %>%
  mutate(
     Date = as.Date(startDateTime)
   )

SoilRadiation <-  SoilRadiation %>% 
  mutate(
    Row = row_number()
  )
```

## Air Temperature

```{r}
#same as soil temp where there were multiple sensors, so averaged by every 30 minute period and got rid of any NA values, did this for both the single and triple air temp

SingleAvgAirTemp <- SingleAirTemp %>% 
  group_by(startDateTime) %>%
  summarise(
    SingleMeanAirTemp = mean(tempSingleMean, na.rm = TRUE),
    .groups = "drop"
  )

TripleAvgAirTemp <- TripleAirTemp %>% 
  group_by(startDateTime) %>%
  summarise(
    TripleMeanAirTemp = mean(tempTripleMean, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
#combined the single and triple aspirated air temps into one df 
ONAQAirTemp <- SingleAvgAirTemp %>% 
  left_join(
    TripleAvgAirTemp %>% select(startDateTime, TripleMeanAirTemp),
    by = "startDateTime"
  )
```

```{r}
#created new column in this df that was the average of the single and triple aspirated air temps
ONAQAirTemp <- ONAQAirTemp %>% 
  mutate(
    AvgAirTemp_C = ((SingleMeanAirTemp + TripleMeanAirTemp)/2)
  )
```

```{r}
#added in more accurate date and time column
ONAQAirTemp <- ONAQAirTemp %>% 
  mutate(
    Date = as.Date(startDateTime))

#adding fahrenheit
ONAQAirTemp <- ONAQAirTemp %>% 
  mutate(
    Temp_F = AvgAirTemp_C * (9/5) + 32)

#adding kelvin
ONAQAirTemp <- ONAQAirTemp %>% 
  mutate(
    Temp_Kelvin = AvgAirTemp_C + 273.15)
#kelvin is used later for other calculations, and added fahrenheit to better understand the data

```

## Utah Lake Isotopes

### From WaterIsotopes.org

```{r}
#renaming some columns for ease and understanding
UtahLakeCounty <- UtahLakeCounty %>%
  rename(
  "Elevation" = "Elevation_mabsl",
  "SD2H" = "d2H_Analytical_SD",
  "SD18O" = "d18O_Analytical_SD",
  "startDateTime" = "Collection_Date"
)

#getting rid of unnecessary columns that are unneccessary
UtahLakeCounty <- UtahLakeCounty %>% 
  select(-Start_Date, -Start_Time_Zone, -Collection_Time_Zone, -Phase, -Depth_meters, -Sample_Comments, -D17O, -D17O_Analytical_SD, -WI_Analysis_Source, -Site_Comments)
```

```{r}
#creating main df, getting rid of sample types named below only leaving lake, river, and spring samples
UtahLakeCounty <- UtahLakeCounty %>% 
  filter(!Type %in% c("Tap", "Milk", "Sprinkler", "Ground"))

```

```{r}
#adding date (20XX-XX)
UtahLakeCounty <- UtahLakeCounty %>%
  mutate(
    Date = as.Date(startDateTime))

#removed because wasn't necessary for analysis but originally had for looking at graphs
#adding Year column
# UtahLakeSamples <- UtahLakeSamples %>% 
#   mutate(
#     Year = format(startDateTime, "%Y")
#   )

#creating DONOTUSE row
UtahLakeCounty <- UtahLakeCounty %>% 
  mutate(
    DONOTUSE = FALSE
  )

```

### From Zanazzi

```{r}

#rename func to make easier and consistent columns
UtahLakeZanazzi <- UtahLakeZanazzi %>% 
  rename(
    "Site_Name" = "...2",
    "Latitude" = "...3",
    "Longitude" = "...4",
    "startDateTime" = "...5",
    "d2H" = "...6",
    "SD2H" = "...7",
    "d18O" = "...8",
    "SD18O" = "...9"
  )

#if d2h is NA, don't include
UtahLakeZanazzi <- UtahLakeZanazzi %>% 
  filter(!is.na(d2H)) 
#gets rid of first row because it was either blank or had column labels
UtahLakeZanazzi <- UtahLakeZanazzi[-c(1), ]

#got rid of these columns because they were werid format issues
UtahLakeZanazzi <- UtahLakeZanazzi %>% select(-`Table S1. Biweekly Samples`)

#turns collection data into a numeric value becasue originally was character 
UtahLakeZanazzi$startDateTime <- as.numeric(UtahLakeZanazzi$startDateTime)

#then turns collection data into data format, excel file did something weird where it was a weird number that represented number of days since dec. 30th 1899 so switched over to actual data format
UtahLakeZanazzi$startDateTime <- as.Date(UtahLakeZanazzi$startDateTime, origin = "1899-12-30")

```

```{r}

#making everything numeric, because everything started as a character value
UtahLakeZanazzi$d18O <- as.numeric(UtahLakeZanazzi$d18O)

UtahLakeZanazzi$d2H <- as.numeric(UtahLakeZanazzi$d2H)

UtahLakeZanazzi$Latitude <- as.numeric(UtahLakeZanazzi$Latitude)

UtahLakeZanazzi$Longitude <- as.numeric(UtahLakeZanazzi$Longitude)

UtahLakeZanazzi$SD2H <- as.numeric(UtahLakeZanazzi$SD2H)

UtahLakeZanazzi$SD18O <- as.numeric(UtahLakeZanazzi$SD18O)


#creating type row, this made it easier to read to the utah county data
UtahLakeZanazzi <- UtahLakeZanazzi %>% 
  mutate(Type = "Zanazzi")

#creating DONOTUSE row
UtahLakeZanazzi <- UtahLakeZanazzi %>%
  mutate(
    DONOTUSE = FALSE
  )

```

### Combining Zanazzi with water isotope

```{r}
#creating combined df, also added on row with row number so could get rid of unwanted values through donotuse and organization purposes
UtahLakeWaterSamples <- bind_rows(UtahLakeCounty, UtahLakeZanazzi) %>% 
  mutate(
    Row = row_number()
  )
```

```{r}
#getting rid of far away sites or other weird #s, from map below 
UtahLakeWaterSamples <- UtahLakeWaterSamples %>% 
      mutate(DONOTUSE = Row %in% c(9, 19, 7, 18, 8, 17, 21 
          , TRUE, DONOTUSE))
```

# Analysis

## Weighted Precipitation

$$\delta_w = \frac{\sum_{i=1}^{n} P_i \delta_i}{\sum_{i=1}^{n} P_i}$$

## Setting up dfs for MC

```{r}
#1 day averages for each continuously measured value
ONAQAtmosd18O_1day <- create_period_summary(
  ONAQAtmos,
  "d18Omean",
  1
)

ONAQAtmosd2H_1day <- create_period_summary(
  ONAQAtmos,
  "d2Hmean",
  1
)

ONAQHumid_1day <- create_period_summary(
  ONAQHumid,
  "RHMean",
  1
)

ONAQAirTempC_1day <- create_period_summary(
  ONAQAirTemp,
  "AvgAirTemp_C",
  1
)

ONAQAirTempKelvin_1day <- create_period_summary(
  ONAQAirTemp,
  "Temp_Kelvin",
  1
)

ONAQSoilRad_1day <- create_period_summary(
  SoilRadiation,
  "difRadMean",
  1
)

#7 day averages of each continuous values - thought this would be more impactful in down the line analysis
ONAQAtmosd18O_7day <- create_period_summary(
  ONAQAtmos,
  "d18Omean",
  7
)

ONAQAtmosd2H_7day <- create_period_summary(
  ONAQAtmos,
  "d2Hmean",
  7
)

ONAQHumid_7day <- create_period_summary(
  ONAQHumid,
  "RHMean",
  7
)

ONAQAirTempC_7day <- create_period_summary(
  ONAQAirTemp,
  "AvgAirTemp_C",
  7
)

ONAQAirTempKelvin_7day <- create_period_summary(
  ONAQAirTemp,
  "Temp_Kelvin",
  7
)

ONAQSoilRad_7day <- create_period_summary(
  SoilRadiation,
  "difRadMean",
  7
)
```

## Creating MC for each variable

```{r}

#running monte carlo on atmospheric oxygen-18 isotopes
MC_dAd18O <- vector(
  mode = "list",
  length = nrow(ONAQAtmosd18O_1day)
)
for(i in 1:nrow(ONAQAtmosd18O_1day)){
  MC_dAd18O[[i]] <- rnorm(
    n = 1000,
    mean = ONAQAtmosd18O_1day$Mean[i],
    sd = ONAQAtmosd18O_1day$SD[i]
  )}


#running monte carlo on atmoshperic hydrogen-2 isotopes
MC_dAd2H <- vector(
  mode = "list",
  length = nrow(ONAQAtmosd2H_1day)
)
for(i in 1:nrow(ONAQAtmosd2H_1day)){
  MC_dAd2H[[i]] <- rnorm(
    n = 1000,
    mean = ONAQAtmosd2H_1day$Mean[i],
    sd = ONAQAtmosd2H_1day$SD[i]
  )}

#monte carlo for umidity measurments
MC_Humid <- vector(
  mode = "list",
  length = nrow(ONAQHumid_1day)
)
for(i in 1:nrow(ONAQHumid_1day)){
  
  MC_Humid[[i]] <- rnorm(
    n = 1000,
    mean = ONAQHumid_1day$Mean[i],
    sd = ONAQHumid_1day$SD[i]
  )}

#monte carlo for air temp in degrees celsius
MC_AirTempC <- vector(
  mode = "list",
  length = nrow(ONAQAirTempC_1day)
)
for(i in 1:nrow(ONAQAirTempC_1day)){
  
  MC_AirTempC[[i]] <- rnorm(
    n = 1000,
    mean = ONAQAirTempC_1day$Mean[i],
    sd = ONAQAirTempC_1day$SD[i]
  )}

#monte carlo for air temp in degrees kelvin
MC_AirTempKelvin <- vector(
  mode = "list",
  length = nrow(ONAQAirTempKelvin_1day)
)
for(i in 1:nrow(ONAQAirTempKelvin_1day)){
  
  MC_AirTempKelvin[[i]] <- rnorm(
    n = 1000,
    mean = ONAQAirTempKelvin_1day$Mean[i],
    sd = ONAQAirTempKelvin_1day$SD[i]
  )}

#monte carlo for soil radiation
MC_SoilRad <- vector(
  mode = "list",
  length = nrow(ONAQSoilRad_1day)
)
for(i in 1:nrow(ONAQSoilRad_1day)){
  
  MC_SoilRad[[i]] <- rnorm(
    n = 1000,
    mean = ONAQSoilRad_1day$Mean[i],
    sd = ONAQSoilRad_1day$SD[i]
  )}


#monte carlo for lake isotopes oxygen 18
MC_UtahLaked18O <- vector(
  mode = "list",
  length = nrow(UtahLakeWaterSamples)
)
for(j in 1:nrow(UtahLakeWaterSamples)){
  
  MC_UtahLaked18O[[j]] <- rnorm(
    n = 1000,
    mean = UtahLakeWaterSamples$d18O[j],
    sd = UtahLakeWaterSamples$SD18O[j]
  )}

#monte carlo for lake isotopes hydrogen 2
MC_UtahLaked2H <- vector(
  mode = "list",
  length = nrow(UtahLakeWaterSamples)
)
for(j in 1:nrow(UtahLakeWaterSamples)){
  
  MC_UtahLaked2H[[j]] <- rnorm(
    n = 1000,
    mean = UtahLakeWaterSamples$d2H[j],
    sd = UtahLakeWaterSamples$SD2H[j]
  )}
```

## Alpha plus

$$
\alpha^+_{d18O}= exp[-\frac{7.685}{10^3}+\frac{6.7123}{(273.15+T)}-\frac{1666.4}{{(273.15+T)}^{2}}+\frac{350410}{{(273.15+T)}^{3}}]
$$

```{r}
MC_alphastar_d18O <- vector("list", length(ONAQAirTempKelvin_1day))

for(i in 1:nrow(ONAQAirTempKelvin_1day)) {

  MC_alphastar_d18O[[i]] <- exp(
    -7.685/10^3 +
    6.7123/(MC_AirTempKelvin[[i]]) -
    1666.4/(MC_AirTempKelvin[[i]]^2) +
    350410/(MC_AirTempKelvin[[i]]^3)
  )

}
```

$$
\alpha^+_{d2H} = exp[1158.8*(\frac{(273.15 + T)^3}{10^{12}}-1620.1(\frac{(273.15+T)^2}{10^9})+794.84(\frac{(273.15+T)}{10^6})-\frac{161.04}{10^3}+\frac{2999200}{(273.15+T)^3}]
$$

```{r}
MC_alphastar_d2H <- vector("list", length(ONAQAirTempKelvin_1day))

for(i in 1:nrow(ONAQAirTempKelvin_1day)) {
  MC_alphastar_d2H[[i]] <- exp(
    1158.8 * ((MC_AirTempKelvin[[i]]^3)/10^12) -
      1620.1 *((MC_AirTempKelvin[[i]]^2)/10^9) +
      794.84 *((MC_AirTempKelvin[[i]])/10^6) - 
      161.04/(10^3) +
      2999200/((MC_AirTempKelvin[[i]])^3)
  )
}
```
